---
title: "16s Analysis Report"
output:
  pdf_document:
    
    
    df_print: kable
  html_document:
    df_print: kable
    smart: false
---
<!-- This is sometimes nessecary, it will manifest with a unable to load shared object '/usr/local/lib/R/ error -->
<!-- you can run ldconfig -p|grep $BROKEN_PACKAGE -->
<!-- then you take the stem of where the bin for that package is and replace it below -->

```{r handle lib errors, echo=FALSE, message=FALSE, warning=FALSE}
Sys.setenv(LD_LIBRARY_PATH = "/lib64")
Sys.setenv(PKG_CONFIG_PATH = "/lib64/pkgconfig")
#Sys.setenv(R_REMOTES_NO_ERRORS_FROM_WARNINGS="true")
if(!require(renv)){install.packages("renv", repos="http://cran.us.r-project.org")}
library(renv)
renv::isolate()
renv::settings$snapshot.type("all")
#renv::init(bioconductor = TRUE)
renv::init()
install.packages("BiocManager")
BiocManager::install(version = "3.12", ask=FALSE)
BiocManager::repositories(version="3.12")
if(!require(rhdf5)){ renv::install("bioc::rhdf5",version="3.12") }
if(!require(Biostrings)) {renv::install("bioc::Biostrings",version="3.12") }
if(!require(Biobase)) {renv::install("bioc::Biobase",version="3.12")}
if(!require(phyloseq)) {renv::install("bioc::phyloseq",version="3.12")}
if(!require(biomformat)) {renv::install("bioc::biomformat",version="3.12")}
renv::restore()
```

```{r function to check how the document is being rendered, message=FALSE, warning=FALSE, include=FALSE}
getOutputFormat <- function() {
  output <- rmarkdown:::parse_yaml_front_matter(
    readLines(knitr::current_input())
    )$output
  if (is.list(output)){
    return(names(output)[1])
  } else {
    return(output[1])
  }
}
```

```{r install/load libraries, message=FALSE, warning=FALSE, include=FALSE}


if(!require(kableExtra)) {renv::install("kableExtra")}
if(!require(remotes)){renv::install("remotes")}
if(!require(ggplot2)) {renv::install("tidyverse/ggplot2@v3.3.2")}
if(!require(tidyverse)) {renv::install("tidyverse@1.3.1")}


if (!require("devtools", quietly = TRUE)){renv::install("devtools")}
if(!require(qiime2R)) {renv::install("jbisanz/qiime2R@d1ad96657ada993cf6c2841b29113a4f635c6b56")} # current version is 0.99.20
if(!require(phyloseq)) {BiocManager::install("phyloseq", version="3.12")}
if(!require(gtools)) {renv::install("gtools")}
if(!require(reshape2)) {renv::install("reshape2")}
if(!require(forcats)) {renv::install("forcats")}
if(!require(gridExtra)) {renv::install("gridExtra")}


if(!require(pheatmap)){renv::install("pheatmap")}
if(!require(rstatix)){renv::install("rstatix@0.7.0")}
#if(!require(knitr)){renv::install("knitr")}
if(!require(tidyverse)){renv::install("tidyverse")}
if(!require(ampvis2)){renv::install("MadsAlbertsen/ampvis2@eaef3f8df7e3fc90ef5c8c7f241f87fa05afb1b8")}
if(!require(devtools)) {renv::install("devtools")}
if(!require(ggdendro)){renv::install("ggdendro@0.1.22", type="source")}
if(!require(vegan)){renv::install("vegan", type="source")}
#if(!require(ape)){renv::install("ape")}
#if(!require(adegenet)){renv::install("adegenet")}
#if(!require(ade4)){renv::install("ade4")}
#if(!require(pegas)){renv::install("pegas@1.0")}

if(!require(ggvegan)){renv::install("gavinsimpson/ggvegan@4bc6ee9945dd9229ed486409c0acab9413b8c9af")}
if(!require(ggpubr)){renv::install("ggpubr")}
if(!require(dplyr)){renv::install("dplyr")}
if(!require(jamba)){ renv::install("jmw86069/jamba@0.0.6.900")}
if(!require(ggConvexHull)){renv::install("cmartin/ggConvexHull@e05904a2ea4747712c0733a85a1e645c308e24ae")}
if(!require(PERMANOVA)){renv::install("PERMANOVA")}
# library(remotes)
library(openssl)
library(gtools)
library(kableExtra)
library(qiime2R)
library(phyloseq)
library(reshape2)
library(forcats)
library(gridExtra)

#library(microbiome)

library(pheatmap)
library(rstatix)
#library(knitr)
#library(tidyverse)
library(reshape2)
library(ampvis2)
library(ggdendro)
library(vegan)
library(pegas)
library(ggvegan)

library(ggpubr)
library(dplyr)
library(jamba)
library(ggConvexHull)
library(PERMANOVA)
library(tinytex)
```
<!-- input files required:  -->
<!-- item_of_interest.csv -->
<!-- table-dada2.qza -->
<!-- rooted-tree.qza -->
<!-- taxonomy.qza -->
<!-- metadata.tsv -->
<!-- image_.*_graph.png[any number of files that match that pattern for phylo trees] -->
<!-- *cutoffs.csv -->
<!-- *dada2_stats.tsv -->
<!-- *sampling_depth.csv -->
<!-- obs/metadata.tsv -->
<!-- shannon/metadata.tsv -->
<!-- simpson/metadata.tsv -->
<!-- chao1/metadata.tsv -->
<!-- ace/metadata.tsv -->
<!-- faith_pd/metadata.tsv -->
<!-- core-metrics-results/bray_curtis_pcoa_results.qza -->
<!-- core-metrics-results/shannon_vector.qza -->
<!-- core-metrics-results/jaccard_pcoa_results.qza -->
<!-- core-metrics-results/unweighted_unifrac_pcoa_results.qza -->
<!-- core-metrics-results/weighted_unifrac_pcoa_results.qza -->
<!-- alpha-rare/observed_otus.csv -->



```{r parse item of interest and make outdir for figures, echo=FALSE, message=FALSE, warning=FALSE}
#item of interest, comes from bash and python script that orchestrates everything
wd <- getwd()
ioi <- read.csv("item_of_interest.csv")
ioi<-as.character(ioi)

ioi_ord <- read.csv("order_item_of_interest.csv")
colnames(ioi_ord) <- ioi
ioi_ord <- as.list(ioi_ord)[[ioi]]

outdir_name <- "Figures"
if(!dir.exists(paste0("./",outdir_name))) {dir.create(paste0("./",outdir_name))}
outdir <- paste0("./",outdir_name)


```

## 1 Species Distribution 

### 1.1 Relative Abundance of Top 10 Phyla

```{r stacked bar chart at Phylum level, echo=FALSE, fig.height=12, fig.width=18, message=FALSE, warning=FALSE}
options(bitmapType='cairo')
level = "Phylum"

#table_dada2 <- "./qiime/table-dada2.qza"
table_dada2 <- "./qiime/core-metric-results/rarefied_table.qza"
rooted_tree <- "./qiime/rooted-tree.qza"
taxonomy_file <- "./qiime/taxonomy.qza"
metadata_file <- "./metadata.tsv"


#Create phyloseq object from qiime objects
cycle_1 <- qza_to_phyloseq(table_dada2,rooted_tree,taxonomy_file,metadata_file)

#Select number of taxa at phylum level to plot
topN <- 10

#selects top n taxa and creates phyloseq object
cycle_n_names <- sort(tapply(taxa_sums(cycle_1), tax_table(cycle_1)[,as.character(level)], sum), TRUE) [1:topN]
cycle_n <- subset_taxa(cycle_1, get(level) %in% names(cycle_n_names))
cycle_n_glom <- tax_glom(cycle_n, taxrank = level)

#selects top n + 1 to the end of table
rest_tax <- sort(tapply(taxa_sums(cycle_1), tax_table(cycle_1)[,as.character(level)], sum), TRUE) [topN+1:length(tax_table(cycle_1))]
rest <- subset_taxa(cycle_1, get(level) %in% names(rest_tax))
rest_glom <- tax_glom(rest, taxrank = level)

#collapse all phyla not in top 10 into one taxon
rest_merge <- merge_taxa(rest, taxa_names(rest))

#create dummy taxonomy classification for collapsed taxa
taxvec <- c("Other","Other","Other","Other","Other", "Other","Other")
taxlist <- list(OTU1=parse_taxonomy_default(taxvec))
tax_tab <- build_tax_table(taxlist)
colnames(tax_tab) <- c("Kingdom","Phylum","Class","Order","Family","Genus","Species")
rownames(tax_tab) <- rownames(tax_table(rest_merge))
tax_table(rest_merge) <- tax_tab

#attaches top n phyloseq object with the other phyloseq object
merge_tax <- merge_phyloseq(tax_table(cycle_n), tax_table(rest_merge))
merge_otu <- merge_phyloseq(otu_table(cycle_n), otu_table(rest_merge))
merge_samp <- merge_phyloseq(sample_data(cycle_n), sample_data(rest_merge))

merged <- merge_phyloseq(merge_tax, merge_otu, merge_samp)

#calculates relative abundance based on sample provided and turns that into a dataframe for ggplot2 plotting
cycle_1_phylum <- merged %>%
  tax_glom(taxrank = level) %>%
  transform_sample_counts(function(x) {x/sum(x)}) %>%
  psmelt() %>%
  arrange(Abundance)


#ggplot2 stacked bar chart separated by sample in the metadata from qiime2
by_sample <- ggplot(cycle_1_phylum, aes(x=Sample, y=Abundance, fill=fct_relevel(fct_reorder(cycle_1_phylum[[level]], Abundance),"Other"))) + 
  geom_bar(stat= "identity") +
  xlab("Sample ID") +
  ylab(label="Relative Abundance") +
  guides(fill=guide_legend(title=level))+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90,vjust=0.5, hjust=1))+
  theme(text=element_text(size = 20))

print(by_sample)
ggsave("stacked_bar_phylum_by_sample.png", plot=by_sample, path=outdir)

#calculates relative abundance based on item of interest provided in original bash script
cy_phylum <- merged %>%
  tax_glom(taxrank = level) %>%
  merge_samples(ioi,fun=sum) %>%
  transform_sample_counts(function(x) {x/sum(x)}) %>%
  psmelt() %>%
  arrange(Abundance)

cy_phylum$Sample <- factor(cy_phylum$Sample, levels=ioi_ord)

#ggplot2 stacked bar chart separated by item of interet in the metadata from qiime2 identified by user
by_treatment <- ggplot(cy_phylum, aes(x=Sample, y=Abundance, fill=fct_relevel(fct_reorder(cy_phylum[[level]],Abundance),"Other"))) + 
  geom_bar(stat="identity") +
  guides(fill=guide_legend(title=level)) + 
  xlab(label = ioi) +
  ylab(label="Relative Abundance") +
  theme_minimal()+
  theme(text=element_text(size = 20))


#prints charts to rnotebook

print(by_treatment)

#saves charts to directory

ggsave("stacked_bar_phylum_by_treatment.png", plot=by_treatment, path=outdir)

```




### 1.2 Relative Abundance of Top 10 Genera

```{r stacked bar chart at Genus level, echo=FALSE, fig.height=12, fig.width=18, message=FALSE, warning=FALSE}
level = "Genus"

#Create phyloseq object from qiime objects
cycle_1 <- qza_to_phyloseq(table_dada2,rooted_tree,taxonomy_file,metadata_file)

#Select number of taxa at genus level to plot
topN <- 10

#selects top n taxa and creates phyloseq object
cycle_n_names <- sort(tapply(taxa_sums(cycle_1), tax_table(cycle_1)[,as.character(level)], sum), TRUE) [1:topN]
cycle_n <- subset_taxa(cycle_1, get(level) %in% names(cycle_n_names))
cycle_n_glom <- tax_glom(cycle_n, taxrank = level)

#selects top n + 1 to the end of table
rest_tax <- sort(tapply(taxa_sums(cycle_1), tax_table(cycle_1)[,as.character(level)], sum), TRUE) [topN+1:length(tax_table(cycle_1))]
rest <- subset_taxa(cycle_1, get(level) %in% names(rest_tax))
rest_glom <- tax_glom(rest, taxrank = level)

#collapse all genera not in top 10 into one taxon
rest_merge <- merge_taxa(rest, taxa_names(rest))

#create dummy taxonomy classificaiton for collapsed taxa
taxvec <- c("Other","Other","Other","Other","Other", "Other","Other")
taxlist <- list(OTU1=parse_taxonomy_default(taxvec))
tax_tab <- build_tax_table(taxlist)
colnames(tax_tab) <- c("Kingdom","Phylum","Class","Order","Family","Genus","Species")
rownames(tax_tab) <- rownames(tax_table(rest_merge))
tax_table(rest_merge) <- tax_tab

#attaches top n phyloseq object with the other phyloseq object
merge_tax <- merge_phyloseq(tax_table(cycle_n), tax_table(rest_merge))
merge_otu <- merge_phyloseq(otu_table(cycle_n), otu_table(rest_merge))
merge_samp <- merge_phyloseq(sample_data(cycle_n), sample_data(rest_merge))

merged <- merge_phyloseq(merge_tax, merge_otu, merge_samp)

#calculate relative abundance based on sample
cycle_1_genus <- merged %>%
  tax_glom(taxrank = level) %>%
  transform_sample_counts(function(x) {x/sum(x)}) %>%
  psmelt() %>%
  arrange(Abundance)

#ggplot of stacked bar chart separated by sample
by_sample <- ggplot(cycle_1_genus, aes(x=Sample, y=Abundance, fill=fct_relevel(fct_reorder(cycle_1_genus[[level]], Abundance),"Other"))) + 
  geom_bar(stat= "identity") +
  xlab("Sample ID") +
  ylab(label="Relative Abundance") +
  guides(fill=guide_legend(title=level))+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90,vjust=0.5, hjust=1)) +
  theme(text=element_text(size = 20))

print(by_sample)
ggsave("stacked_bar_genus_by_sample.png", plot=by_sample, path=outdir)

#calculate relative abundance based on treatment 
cy_genus <- merged %>%
  tax_glom(taxrank = level) %>%
  merge_samples(ioi,fun=sum) %>%
  transform_sample_counts(function(x) {x/sum(x)}) %>%
  psmelt() %>%
  arrange(Abundance)

cy_genus$Sample <- factor(cy_genus$Sample, levels=ioi_ord)

#ggplot figure of stacked bar chart separated by treatment
by_treatment <- ggplot(cy_genus, aes(x=Sample, y=Abundance, fill=fct_relevel(fct_reorder(cy_genus[[level]],Abundance),"Other"))) + 
  geom_bar(stat="identity") +
  guides(fill=guide_legend(title=level)) + 
  xlab(label = ioi) +
  ylab(label="Relative Abundance") +
  theme_minimal()+
  theme(text=element_text(size = 20))

#prints charts for rnotebook

print(by_treatment)

#saves copies of figures to outdir

ggsave("stacked_bar_genus_by_treatment.png", plot = by_treatment, path=outdir)

```




### 1.3 Relative Abundance of Top 10 Species

```{r stacked bar chart at Species level, echo=FALSE, fig.height=12, fig.width=18, message=FALSE, warning=FALSE}
level = "Species"

#Create phyloseq object from qiime objects
cycle_1 <- qza_to_phyloseq(table_dada2,rooted_tree,taxonomy_file,metadata_file)

#Select number of taxa at species level to plot
topN <- 10

#selects top n taxa and creates phyloseq object
cycle_n_names <- sort(tapply(taxa_sums(cycle_1), tax_table(cycle_1)[,as.character(level)], sum), TRUE) [1:topN]
cycle_n <- subset_taxa(cycle_1, get(level) %in% names(cycle_n_names))
cycle_n_glom <- tax_glom(cycle_n, taxrank = level)

#selects top n + 1 to the end of table
rest_tax <- sort(tapply(taxa_sums(cycle_1), tax_table(cycle_1)[,as.character(level)], sum), TRUE) [topN+1:length(tax_table(cycle_1))]
rest <- subset_taxa(cycle_1, get(level) %in% names(rest_tax))
rest_glom <- tax_glom(rest, taxrank = level)

#collapse all species not in top 10 into one taxon
rest_merge <- merge_taxa(rest, taxa_names(rest))

#create dummy taxonomy classificaiton for collapsed taxa
taxvec <- c("Other","Other","Other","Other","Other", "Other","Other")
taxlist <- list(OTU1=parse_taxonomy_default(taxvec))
tax_tab <- build_tax_table(taxlist)
colnames(tax_tab) <- c("Kingdom","Phylum","Class","Order","Family","Genus","Species")
rownames(tax_tab) <- rownames(tax_table(rest_merge))
tax_table(rest_merge) <- tax_tab

#attaches top n phyloseq object with the other phyloseq object
merge_tax <- merge_phyloseq(tax_table(cycle_n), tax_table(rest_merge))
merge_otu <- merge_phyloseq(otu_table(cycle_n), otu_table(rest_merge))
merge_samp <- merge_phyloseq(sample_data(cycle_n), sample_data(rest_merge))

merged <- merge_phyloseq(merge_tax, merge_otu, merge_samp)

#calculates relative abundance based on sample
cycle_1_species <- merged %>%
  tax_glom(taxrank = level) %>%
  transform_sample_counts(function(x) {x/sum(x)}) %>%
  psmelt() %>%
  arrange(Abundance)

#ggplot stacked bar chart separated by species
by_sample <- ggplot(cycle_1_species, aes(x=Sample, y=Abundance, fill=fct_relevel(fct_reorder(cycle_1_species[[level]], Abundance),"Other"))) + 
  geom_bar(stat= "identity") +
  xlab("Sample ID") +
  ylab(label="Relative Abundance") +
  guides(fill=guide_legend(title=level))+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90,vjust=0.5, hjust=1)) +
  theme(text=element_text(size = 20))

print(by_sample)
ggsave("stacked_bar_species_by_sample.png", plot=by_sample, path=outdir)


#calculates relative abundance based on treatment
cy_species <- merged %>%
  tax_glom(taxrank = level) %>%
  merge_samples(ioi,fun=sum) %>%
  transform_sample_counts(function(x) {x/sum(x)}) %>%
  psmelt() %>%
  arrange(Abundance)

cy_species$Sample <- factor(cy_species$Sample, levels=ioi_ord)

#ggplot stacked bar chart separated by treatment
by_treatment <- ggplot(cy_species, aes(x=Sample, y=Abundance, fill=fct_relevel(fct_reorder(cy_species[[level]],Abundance),"Other"))) + 
  geom_bar(stat="identity") +
  ylab(label="Relative Abundance") +
  guides(fill=guide_legend(title=level)) + 
  xlab(label = ioi) +
  theme_minimal()+
  theme(text=element_text(size = 20))

#prints figures to rnotebook

print(by_treatment)

#saves copies of figures to directory

ggsave("stacked_bar_species_by_treatment.png", plot=by_treatment, path=outdir)

```




## 2 Phylogentic Trees

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# if(getOutputFormat() =='pdf_document'){
#   print("see end of report")
# }

```


```{r phylo-tree", echo=FALSE, message=FALSE, warning=FALSE, fig.pos='h', out.width="100%", fig.cap="\\newline{}"}
#if(getOutputFormat() == 'html_document') {
  library(knitr)
  tree_filenames <- list.files("./graphlan/phylo_trees",pattern="image_.*_graph.png",full.names=T)
  tree_filenames <- mixedsort(tree_filenames)

  include_graphics(tree_filenames)
  
#} 
# 
# if(getOutputFormat() =='pdf_document'){
#   
#   print("see end of report")
#   tree_filenames <- list.files("./graphlan/phylo_trees",pattern="image_.*_pdf_g.png", full.names=T)
#   tree_filenames <- mixedsort(tree_filenames)
# 
#   include_graphics(tree_filenames)
#   
# }
```




## 3 Species Abundance Heatmap

```{r heatmap at the Genus level annotatted with item of interest, echo=FALSE, fig.height=10, fig.width=14, message=FALSE, warning=FALSE}
devtools::source_gist("8d0ca4206a66be7ff6d76fc4ab8e66c6")
cycle_1 <- qza_to_phyloseq(table_dada2,rooted_tree,taxonomy_file,metadata_file) 
cycle_1@sam_data[[ioi]] <- factor(cycle_1@sam_data[[ioi]], levels= ioi_ord)

cycle_1_amp <- phyloseq_to_ampvis2(cycle_1)

heatmap <- ampvis2::amp_heatmap(cycle_1_amp,
                     group_by = "SampleID", 
                     tax_aggregate = "Genus",
                     tax_show = 30,
                     facet_by = ioi,
                     tax_add = "Phylum",
                     plot_values = FALSE,
                     plot_colorscale = "log10") +
  theme(text=element_text(size = 20))

heatmap
ggsave("heatmap.png", plot=heatmap, path=outdir)

```


```{r read in cutoffs used by qiime, message=FALSE, warning=FALSE, include=FALSE}
cutoffs = read.csv("./qiime/cutoffs.csv",sep = ',',strip.white = TRUE)
#cutoffs = knitr::kable(cutoffs)
cutoffs %>% kbl() %>% kable_styling(bootstrap_options = c("striped","condensed","responsive"))
```


```{r make data table for dada2 stats from qiime, message=FALSE, warning=FALSE, include=FALSE}
dadastat = read.csv("./dada2_stats.tsv",sep = '\t',strip.white = TRUE)
#dadastat_t = dadastat %>% kbl() %>% kable_classic(full_width = F)
dadastat %>% kbl() %>% kable_styling(bootstrap_options = c("striped","condensed","responsive")) 
```

```{r table for sampling depth in qiime, message=FALSE, warning=FALSE, include=FALSE}
depth = read.csv("./qiime/sampling_depth.csv", sep=",")
depth %>% kbl() %>% kable_styling(bootstrap_options = c("striped","condensed","responsive"))
```




## 4 Alpha Diversity Table

```{r echo=FALSE, message=FALSE, warning=FALSE}
obs <- read.table("./qiime/obs/metadata.tsv", header = TRUE)
shannon <- read.table("./qiime/shannon/metadata.tsv", header = TRUE)
simpson <- read.table("./qiime/simpson/metadata.tsv", header=TRUE)
chao1 <- read.table("./qiime/chao1/metadata.tsv", header=TRUE)
ace <- read.table("./qiime/ace/metadata.tsv", header=TRUE)
faith_pd <- read.table("./qiime/faith_pd/metadata.tsv", header=TRUE)

obs <- data.frame(obs[c('id','observed_features')])
shannon <- data.frame(shannon[c('id','shannon_entropy')])
simpson <- data.frame(simpson[c('id','simpson')])
chao1 <- data.frame(chao1[c('id','chao1')])
ace <- data.frame(ace[c("id","ace")])
faith_pd <- data.frame(faith_pd[c("id","faith_pd")])

alpha_summary <- Reduce(function(...) merge(..., by='id', all.x=TRUE), list(obs,shannon,simpson,chao1,ace,faith_pd))
kable(alpha_summary) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```




## 5 Alpha Diversity Plot 




### 5.1 Observed Features 

```{r alpha diversity plot of obs, results = "asis", echo=FALSE, message=FALSE, warning=FALSE}
#subsets and plots observed features in a boxplot format
ioi <- read.csv("item_of_interest.csv")
ioi<-as.character(ioi)
obs_vals = read.csv("./qiime/obs/metadata.tsv", sep='\t')
dims = dim(obs_vals)
obs_vals <- data.frame(obs_vals[2:dims[1],])
obs_vals[[ioi]] <- factor(obs_vals[[ioi]], levels=ioi_ord)
obs_vals$observed_features <- as.numeric(as.character(obs_vals$observed_features))
obs_boxplot <- ggplot(obs_vals, aes_string(x=ioi,y = obs_vals$observed_features, fill=ioi)) + geom_boxplot() + ylab("Observed Features") +theme_bw()
obs_boxplot 

ggsave("obs_boxplot.png", plot=obs_boxplot, path=outdir)

#plots obs values in a violin plot format 
p <- ggplot(obs_vals, aes(x=obs_vals[[ioi]], y=observed_features)) + geom_violin()+ geom_boxplot(width=0.1)
p<- p + stat_summary(fun.y=mean, geom="point", shape=23, size=2)
p<-p + stat_summary(fun.y=median, geom="point", size=2, color="red")
p <- p+geom_jitter(shape=16, position=position_jitter(0.2)) + xlab(ioi) + ylab("Observed Features") + theme_bw()
p

ggsave("obs_violin_plot.png", plot=p, path=outdir)
```


```{r Kruskal-Wallace test for signifiance observed feat,results = "asis", echo=FALSE, message=FALSE, warning=FALSE}
library("ggpubr")
obs <- read.table("./qiime/obs/metadata.tsv", header = TRUE)
obs_stat <- data.frame(obs[c('id',ioi,'observed_features')])

obs_vals[[ioi]] <- factor(obs_vals[[ioi]], levels=ioi_ord)

form <- as.formula(paste0("observed_features ~ ", ioi))
kruskal_table <- rstatix::kruskal_test(data = obs_stat, formula = form)
kruskal_table <- add_significance(kruskal_table, p.col="p",output.col = "p.signif",symbols =c("****", "***", "**", "$*$", "ns"))

kable(kruskal_table) %>% 
  kable_styling(bootstrap_options = c("hover", "condensed", "responsive"))

form = as.formula(paste0("observed_features ~", ioi))

dunn_table <- rstatix::dunn_test(data = obs_stat, formula = form, p.adjust.method = "holm") 
dunn_table<- add_significance(dunn_table, p.col="p.adj",output.col = "p.adj.signif",symbols =c("****", "***", "**", "$*$", "ns"))
kable(dunn_table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

dunn_table <- dunn_table %>% add_xy_position()

p <- ggplot(obs_vals, aes(x=obs_vals[[ioi]], y=observed_features)) + geom_violin()+ geom_boxplot(width=0.1)
p<- p + stat_summary(fun.y=mean, geom="point", shape=23, size=2)
p<-p + stat_summary(fun.y=median, geom="point", size=2, color="red")
p <- p+geom_jitter(shape=16, position=position_jitter(0.2)) + xlab(ioi) + ylab("Observed Features") + theme_bw()
dunn_table<- add_significance(dunn_table, p.col="p.adj",output.col = "p.adj.signif")
p <- p + stat_pvalue_manual(dunn_table, label = "p.adj.signif",hide.ns = TRUE )
p
ggsave("obs_boxplot_w_signif.png", plot=p, path=outdir)

```




### 5.2 Shannon

```{r alpha diversity plot of shannon diversity, echo=FALSE, message=FALSE, warning=FALSE}
shannon_vals <- read.table("./qiime/shannon/metadata.tsv",sep = '\t', header = TRUE)
dims = dim(shannon_vals)
shannon_vals <- data.frame(shannon_vals[2:dims[1],])
shannon_vals[[ioi]] <- factor(shannon_vals[[ioi]], levels=ioi_ord)
shannon_vals$shannon_entropy <- as.numeric(as.character(shannon_vals$shannon_entropy))
shannon_boxplot <- ggplot(shannon_vals, aes_string(x=ioi, y=shannon_vals$shannon_entropy, fill=ioi)) + geom_boxplot() + ylab("Shannon") + theme_bw() + xlab(ioi)
shannon_boxplot

ggsave("shannon_boxplot.png", plot=shannon_boxplot, path=outdir)

p <- ggplot(shannon_vals, aes(x=shannon_vals[[ioi]], y=shannon_entropy)) + geom_violin()+ geom_boxplot(width=0.1)
p<- p + stat_summary(fun.y=mean, geom="point", shape=23, size=2)
p<-p + stat_summary(fun.y=median, geom="point", size=2, color="red")
p <- p+geom_jitter(shape=16, position=position_jitter(0.2)) + xlab(ioi) + ylab("Shannon") + theme_bw()
p

ggsave("shannon_violin_plot.png", plot=p, path=outdir)
```

```{r Kuskal-Wallace test for significance shannon,results = "asis", echo=FALSE, message=FALSE, warning=FALSE}
shannon <- read.table("./qiime/shannon/metadata.tsv", header = TRUE, sep='\t')
shannon_stat <- data.frame(shannon[c('id',ioi,'shannon_entropy')])

shannon_vals[[ioi]] <- factor(shannon_vals[[ioi]], levels=ioi_ord)

form <- as.formula(paste0("shannon_entropy ~ ", ioi))
kruskal_table <- rstatix::kruskal_test(data = shannon_stat, formula =form)
kruskal_table <- add_significance(kruskal_table, p.col="p",output.col = "p.signif",symbols =c("****", "***", "**", "$*$", "ns"))
kable(kruskal_table) %>% 
  kable_styling(bootstrap_options = c("hover", "condensed", "responsive"))

form = as.formula(paste0("shannon_entropy ~", ioi))

dunn_table <- rstatix::dunn_test(data = shannon_stat, formula = form, p.adjust.method = "holm")
dunn_table<- add_significance(dunn_table, p.col="p.adj",output.col = "p.adj.signif",symbols =c("****", "***", "**", "$*$", "ns"))
kable(dunn_table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

dunn_table <- dunn_table  %>% add_xy_position()

p <- ggplot(shannon_vals, aes(x=shannon_vals[[ioi]], y=shannon_entropy)) + geom_violin()+ geom_boxplot(width=0.1)
p<- p + stat_summary(fun.y=mean, geom="point", shape=23, size=2)
p<-p + stat_summary(fun.y=median, geom="point", size=2, color="red")
p <- p+geom_jitter(shape=16, position=position_jitter(0.2)) + xlab(ioi) + ylab("Shannon") + theme_bw()
dunn_table<- add_significance(dunn_table, p.col="p.adj",output.col = "p.adj.signif")
p <- p + stat_pvalue_manual(dunn_table, label = "p.adj.signif",hide.ns = TRUE)
p
ggsave("shannon_boxplot_w_signif.png", plot=p, path=outdir)
```




### 5.3 Simpson

```{r alpha diversity plot of simpson diversity, echo=FALSE, message=FALSE, warning=FALSE}
simpson_vals <- read.table("./qiime/simpson/metadata.tsv",sep = '\t', header = TRUE)
dims = dim(simpson_vals)
simpson_vals <- data.frame(simpson_vals[2:dims[1],])
simpson_vals[[ioi]] <- factor(simpson_vals[[ioi]],levels=ioi_ord)
simpson_vals$simpson <- as.numeric(as.character(simpson_vals$simpson))
simpson_boxplot <- ggplot(simpson_vals, aes_string(x=simpson_vals[[ioi]], y=simpson_vals$simpson, fill=ioi)) + geom_boxplot() + ylab("Simpson") + theme_bw() + xlab(ioi)
simpson_boxplot

ggsave("simpson_boxplot.png", plot=simpson_boxplot, path=outdir)

p <- ggplot(simpson_vals, aes(x=simpson_vals[[ioi]], y=simpson)) + geom_violin()+ geom_boxplot(width=0.1)
p<- p + stat_summary(fun.y=mean, geom="point", shape=23, size=2)
p<-p + stat_summary(fun.y=median, geom="point", size=2, color="red")
p <- p+geom_jitter(shape=16, position=position_jitter(0.2)) + xlab(ioi) + ylab("Simpson") + theme_bw()
p

ggsave("simpson_violin_plot.png", plot=p, path=outdir)
```

```{r Kuskal-Wallace test for significance simpson,results = "asis", echo=FALSE, message=FALSE, warning=FALSE}
simpson <- read.table("./qiime/simpson/metadata.tsv", header = TRUE, sep='\t')
simpson_stat <- data.frame(simpson[c('id',ioi,'simpson')])

simpson_vals[[ioi]] <- factor(simpson_vals[[ioi]], levels=ioi_ord)

form <- as.formula(paste0("simpson ~ ", ioi))
kruskal_table <- rstatix::kruskal_test(data = simpson_stat, formula = form)
kruskal_table <- add_significance(kruskal_table, p.col="p",output.col = "p.signif",symbols =c("****", "***", "**", "$*$", "ns"))
kable(kruskal_table) %>% 
  kable_styling(bootstrap_options = c("hover", "condensed", "responsive"))

form = as.formula(paste0("simpson ~", ioi))

dunn_table <- rstatix::dunn_test(data = simpson_stat, formula = form, p.adjust.method = "holm") 
dunn_table<- add_significance(dunn_table, p.col="p.adj",output.col = "p.adj.signif",symbols =c("****", "***", "**", "$*$", "ns"))
kable(dunn_table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

dunn_table <- dunn_table %>% add_xy_position()

p <- ggplot(simpson_vals, aes(x=simpson_vals[[ioi]], y=simpson)) + geom_violin()+ geom_boxplot(width=0.1)
p<- p + stat_summary(fun.y=mean, geom="point", shape=23, size=2)
p<-p + stat_summary(fun.y=median, geom="point", size=2, color="red")
p <- p+geom_jitter(shape=16, position=position_jitter(0.2)) + xlab(ioi) + ylab("Simpson") + theme_bw()
dunn_table<- add_significance(dunn_table, p.col="p.adj",output.col = "p.adj.signif")
p <- p+ stat_pvalue_manual(dunn_table, label = "p.adj.signif", hide.ns = TRUE)
p
ggsave("simpson_boxplot_w_signif.png", plot=p, path=outdir)
```




### 5.4 Chao 1

```{r alpha diversity plot of Chao 1 diversity, echo=FALSE, message=FALSE, warning=FALSE}
chao1_vals <- read.table("./qiime/chao1/metadata.tsv",sep = '\t', header = TRUE)
dims = dim(chao1_vals)
chao1_vals <- data.frame(chao1_vals[2:dims[1],])
chao1_vals[[ioi]] <- factor(chao1_vals[[ioi]], levels=ioi_ord)
chao1_vals$chao1 <- as.numeric(as.character(chao1_vals$chao1))
chao1_boxplot <- ggplot(chao1_vals, aes_string(x=chao1_vals[[ioi]], y=chao1_vals$chao1, fill=ioi)) + geom_boxplot() + ylab("Chao 1") + theme_bw() + xlab(ioi)
chao1_boxplot

ggsave("chao1_boxplot.png", plot=chao1_boxplot, path=outdir)

p <- ggplot(chao1_vals, aes(x=chao1_vals[[ioi]], y=chao1)) + geom_violin()+ geom_boxplot(width=0.1)
p<- p + stat_summary(fun.y=mean, geom="point", shape=23, size=2)
p<-p + stat_summary(fun.y=median, geom="point", size=2, color="red")
p <- p+geom_jitter(shape=16, position=position_jitter(0.2)) + xlab(ioi) + ylab("Chao 1") + theme_bw()
p

ggsave("chao1_violin_plot.png", plot=p, path=outdir)
```

```{r Kuskal-Wallace test for significance chao1,results = "asis", echo=FALSE, message=FALSE, warning=FALSE}
chao1 <- read.table("./qiime/chao1/metadata.tsv", header = TRUE, sep='\t')
chao1_stat <- data.frame(chao1[c('id',ioi,'chao1')])

chao1_vals[[ioi]] <- factor(chao1_vals[[ioi]], levels=ioi_ord)

form <- as.formula(paste0("chao1 ~ ", ioi))
kruskal_table <- rstatix::kruskal_test(data = chao1_stat , formula = form)
kruskal_table <- add_significance(kruskal_table, p.col="p",output.col = "p.signif",symbols =c("****", "***", "**", "$*$", "ns"))
kable(kruskal_table) %>% 
  kable_styling(bootstrap_options = c("hover", "condensed", "responsive"))

form = as.formula(paste0("chao1 ~", ioi))

dunn_table <- rstatix::dunn_test(data = chao1_stat, formula = form, p.adjust.method = "holm") 
dunn_table<- add_significance(dunn_table, p.col="p.adj",output.col = "p.adj.signif",symbols =c("****", "***", "**", "$*$", "ns"))
kable(dunn_table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

dunn_table <- dunn_table %>% add_xy_position()

p <- ggplot(chao1_vals, aes(x=chao1_vals[[ioi]], y=chao1)) + geom_violin()+ geom_boxplot(width=0.1)
p<- p + stat_summary(fun.y=mean, geom="point", shape=23, size=2)
p<-p + stat_summary(fun.y=median, geom="point", size=2, color="red")
p <- p+geom_jitter(shape=16, position=position_jitter(0.2)) + xlab(ioi) + ylab("Chao 1") + theme_bw()
dunn_table<- add_significance(dunn_table, p.col="p.adj",output.col = "p.adj.signif")
p <- p+stat_pvalue_manual(dunn_table, label="p.adj.signif", hide.ns = TRUE)
p
ggsave("chao1_boxplot_w_signif.png", plot=p, path=outdir)
```




### 5.5 ACE

```{r alpha diversity plot of ACE diversity, echo=FALSE, message=FALSE, warning=FALSE}
ace_vals <- read.table("./qiime/ace/metadata.tsv",sep = '\t', header = TRUE)
dims = dim(ace_vals)
ace_vals <- data.frame(ace_vals[2:dims[1],])
ace_vals[[ioi]] <- factor(ace_vals[[ioi]], levels=ioi_ord)
ace_vals$ace <- as.numeric(as.character(ace_vals$ace))
ace_boxplot <- ggplot(ace_vals, aes_string(x=ace_vals[[ioi]], y=ace_vals$ace, fill=ioi)) + geom_boxplot() + ylab("ACE") + theme_bw() + xlab(ioi)
ace_boxplot

ggsave("ace_boxplot.png", plot=ace_boxplot, path=outdir)

p <- ggplot(ace_vals, aes(x=ace_vals[[ioi]], y=ace)) + geom_violin()+ geom_boxplot(width=0.1)
p<- p + stat_summary(fun.y=mean, geom="point", shape=23, size=2)
p<-p + stat_summary(fun.y=median, geom="point", size=2, color="red")
p <- p+geom_jitter(shape=16, position=position_jitter(0.2)) + xlab(ioi) + ylab("ACE") + theme_bw()
p

ggsave("ace_violin_plot.png", plot=p, path=outdir)
```

```{r Kuskal-Wallace test for significance ace,results = "asis", echo=FALSE, message=FALSE, warning=FALSE}
ace <- read.table("./qiime/ace/metadata.tsv", header = TRUE, sep='\t')
ace_stat <- data.frame(ace[c('id',ioi,'ace')])

ace_vals[[ioi]] <- factor(ace_vals[[ioi]], levels=ioi_ord)

form <- as.formula(paste0("ace ~ ", ioi))
kruskal_table <- rstatix::kruskal_test(data = ace_stat, formula = form)
kruskal_table <- add_significance(kruskal_table, p.col="p",output.col = "p.signif",symbols =c("****", "***", "**", "$*$", "ns"))
kable(kruskal_table) %>% 
  kable_styling(bootstrap_options = c("hover", "condensed", "responsive"))

form = as.formula(paste0("ace ~", ioi))

dunn_table <- rstatix::dunn_test(data = ace_stat, formula = form, p.adjust.method = "holm") 
dunn_table<- add_significance(dunn_table, p.col="p.adj",output.col = "p.adj.signif",symbols =c("****", "***", "**", "$*$", "ns"))
kable(dunn_table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

dunn_table <- dunn_table %>% add_xy_position()

p <- ggplot(ace_vals, aes(x=ace_vals[[ioi]], y=ace)) + geom_violin()+ geom_boxplot(width=0.1)
p<- p + stat_summary(fun.y=mean, geom="point", shape=23, size=2)
p<-p + stat_summary(fun.y=median, geom="point", size=2, color="red")
p <- p+geom_jitter(shape=16, position=position_jitter(0.2)) + xlab(ioi) + ylab("ACE") + theme_bw()
dunn_table<- add_significance(dunn_table, p.col="p.adj",output.col = "p.adj.signif")
p <- p+ stat_pvalue_manual(dunn_table, label = "p.adj.signif", hide.ns = TRUE)
p
ggsave("ace_boxplot_w_signif.png", plot=p, path=outdir)
```




### 5.6 Phylogenetic Distance Over The whole Tree

```{r alpha diversity plot of PD_whole_tree, echo=FALSE, message=FALSE, warning=FALSE}
faith_pd_vals <- read.table("./qiime/faith_pd/metadata.tsv",sep = '\t', header = TRUE)
dims = dim(faith_pd_vals)
faith_pd_vals <- data.frame(faith_pd_vals[2:dims[1],])
faith_pd_vals[[ioi]] <- factor(faith_pd_vals[[ioi]], levels=ioi_ord)
faith_pd_vals$faith_pd <- as.numeric(as.character(faith_pd_vals$faith_pd))
faith_pd_boxplot <- ggplot(faith_pd_vals, aes_string(x=faith_pd_vals[[ioi]], y=faith_pd_vals$faith_pd, fill=ioi)) + geom_boxplot() + ylab("Faith PD") + theme_bw() + xlab(ioi)
faith_pd_boxplot

ggsave("faith_pd_boxplot.png", plot=faith_pd_boxplot, path=outdir)

p <- ggplot(faith_pd_vals, aes(x=faith_pd_vals[[ioi]], y=faith_pd)) + geom_violin()+ geom_boxplot(width=0.1)
p<- p + stat_summary(fun.y=mean, geom="point", shape=23, size=2)
p<-p + stat_summary(fun.y=median, geom="point", size=2, color="red")
p <- p+geom_jitter(shape=16, position=position_jitter(0.2)) + xlab(ioi) + ylab("Faith PD") + theme_bw()
p

ggsave("faith_pd_violin_plot.png", plot=p, path=outdir)
```

```{r Kuskal-Wallace test for significance faith pd,results = "asis", echo=FALSE, message=FALSE, warning=FALSE}
faith_pd <- read.table("./qiime/faith_pd/metadata.tsv", header = TRUE, sep='\t')
faith_pd_stat <- data.frame(faith_pd[c('id',ioi,'faith_pd')])

faith_pd_vals[[ioi]] <- factor(faith_pd_vals[[ioi]], levels=ioi_ord)

form <- as.formula(paste0("faith_pd ~ ", ioi))
kruskal_table <- rstatix::kruskal_test(data = faith_pd_stat, formula = form)
kruskal_table <- add_significance(kruskal_table, p.col="p",output.col = "p.signif",symbols =c("****", "***", "**", "$*$", "ns"))
kable(kruskal_table) %>% 
  kable_styling(bootstrap_options = c("hover", "condensed", "responsive"))

form = as.formula(paste0("faith_pd ~", ioi))

dunn_table <- rstatix::dunn_test(data = faith_pd_stat, formula = form, p.adjust.method = "holm") 
dunn_table<- add_significance(dunn_table, p.col="p.adj",output.col = "p.adj.signif",symbols =c("****", "***", "**", "$*$", "ns"))
kable(dunn_table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

dunn_table <- dunn_table %>% add_xy_position()

p <- ggplot(faith_pd_vals, aes(x=faith_pd_vals[[ioi]], y=faith_pd)) + geom_violin()+ geom_boxplot(width=0.1)
p<- p + stat_summary(fun.y=mean, geom="point", shape=23, size=2)
p<-p + stat_summary(fun.y=median, geom="point", size=2, color="red")
p <- p+geom_jitter(shape=16, position=position_jitter(0.2)) + xlab(ioi) + ylab("Faith PD") + theme_bw()
dunn_table<- add_significance(dunn_table, p.col="p.adj",output.col = "p.adj.signif")
p <- p+stat_pvalue_manual(dunn_table, label="p.adj.signif", hide.ns = TRUE)
p
ggsave("faith_pd_boxplot_w_signif.png", plot=p, path=outdir)
```




## 6 Beta diversity 

### 6.1 Bray curtis ordination

```{r beta diversity plot of bray curtis measurement, echo=FALSE, message=FALSE, warning=FALSE}
library(qiime2R)
metadata<-read_q2metadata(metadata_file)
metadata[[ioi]] <- factor(metadata[[ioi]], levels=ioi_ord)
bray_pcoa <- qiime2R::read_qza("./qiime/core-metric-results/bray_curtis_pcoa_results.qza")
shannon<-qiime2R::read_qza("./qiime/core-metric-results/shannon_vector.qza")
shannon <- shannon$data %>% rownames_to_column("SampleID")


bray_curtis <- bray_pcoa$data$Vectors %>%
  dplyr::select(SampleID, PC1, PC2) %>%
  left_join(metadata) %>%
  left_join(shannon) %>%
  ggplot(aes_string(x="PC1", y="PC2", color=ioi))+
  geom_point() +
  geom_convexhull(alpha=0.5, aes_string(fill=ioi)) +
  theme_q2r() +
  scale_color_discrete(name=ioi)

bray_curtis_no_poly <- bray_pcoa$data$Vectors %>%
  dplyr::select(SampleID, PC1, PC2) %>%
  left_join(metadata) %>%
  left_join(shannon) %>%
  ggplot(aes_string(x="PC1", y="PC2", color=ioi))+
  geom_point() +
  theme_q2r() +
  scale_color_discrete(name=ioi)

bray_curtis
ggsave("bray_curtis_pcoa_ordination.png", plot=bray_curtis, path=outdir)

bray_curtis_no_poly
ggsave("bray_curtis_pcoa_ordination_no_polygon.png", plot=bray_curtis_no_poly, path=outdir)

```




### 6.2 Jaccard ordination

```{r pcoa using Jaccard distance, echo=FALSE, message=FALSE, warning=FALSE}
metadata<-read_q2metadata(metadata_file)
jaccard <- qiime2R::read_qza("./qiime/core-metric-results/jaccard_pcoa_results.qza")
shannon<-qiime2R::read_qza("./qiime/core-metric-results/shannon_vector.qza")
shannon <- shannon$data %>% rownames_to_column("SampleID")

metadata[[ioi]] <- factor(metadata[[ioi]], levels=ioi_ord)

jaccard_pcoa <- jaccard$data$Vectors %>%
  dplyr::select(SampleID, PC1, PC2) %>%
  left_join(metadata) %>%
  left_join(shannon) %>%
  ggplot(aes_string(x="PC1", y="PC2", color=ioi))+
  geom_point() +
  geom_convexhull(alpha=0.5, aes_string(fill=ioi)) +
  theme_q2r() +
  scale_size_continuous(name="Shannon Diversity") +
  scale_color_discrete(name=ioi)

jaccard_pcoa_no_polygon <- jaccard$data$Vectors %>%
  dplyr::select(SampleID, PC1, PC2) %>%
  left_join(metadata) %>%
  left_join(shannon) %>%
  ggplot(aes_string(x="PC1", y="PC2", color=ioi))+
  geom_point() +
  theme_q2r() +
  scale_size_continuous(name="Shannon Diversity") +
  scale_color_discrete(name=ioi)

jaccard_pcoa

ggsave("jaccard_pcoa_ordination.png", plot=jaccard_pcoa, path=outdir)

jaccard_pcoa_no_polygon
ggsave("jaccard_pcoa_ordination_no_polygons.png", plot=jaccard_pcoa_no_polygon, path=outdir)


```




### 6.3 Unweighted unifrac ordination

```{r unweighted unifrac pcoa ordination, echo=FALSE, message=FALSE, warning=FALSE}
metadata<-read_q2metadata(metadata_file)
metadata[[ioi]] <- factor(metadata[[ioi]], levels=ioi_ord)
jaccard <- qiime2R::read_qza("./qiime/core-metric-results/unweighted_unifrac_pcoa_results.qza")
shannon<-qiime2R::read_qza("./qiime/core-metric-results/shannon_vector.qza")
shannon <- shannon$data %>% rownames_to_column("SampleID")


unweight_unifrac_pcoa <- jaccard$data$Vectors %>%
  dplyr::select(SampleID, PC1, PC2) %>%
  left_join(metadata) %>%
  left_join(shannon) %>%
  ggplot(aes_string(x="PC1", y="PC2", color=ioi))+
  geom_point() +
  geom_convexhull(alpha=0.5, aes_string(fill=ioi)) +
  theme_q2r() +
  scale_color_discrete(name=ioi)

unweight_unifrac_pcoa 

ggsave("unweighted_unifrac_pcoa_ordination.png", plot=unweight_unifrac_pcoa, path=outdir)

unweight_unifrac_pcoa_no_polygon <- jaccard$data$Vectors %>%
  dplyr::select(SampleID, PC1, PC2) %>%
  left_join(metadata) %>%
  left_join(shannon) %>%
  ggplot(aes_string(x="PC1", y="PC2", color=ioi))+
  geom_point() +
  theme_q2r() +
  scale_color_discrete(name=ioi)

unweight_unifrac_pcoa_no_polygon

ggsave("unweighted_unifrac_pcoa_ordination_no_polygons.png", plot=unweight_unifrac_pcoa_no_polygon, path=outdir)


```




### 6.4 Weighted unifrac ordination

```{r echo=FALSE, message=FALSE, warning=FALSE}
metadata<-read_q2metadata(metadata_file)
metadata[[ioi]] <- factor(metadata[[ioi]], levels=ioi_ord)
jaccard <- qiime2R::read_qza("./qiime/core-metric-results/weighted_unifrac_pcoa_results.qza")
shannon<-qiime2R::read_qza("./qiime/core-metric-results/shannon_vector.qza")
shannon <- shannon$data %>% rownames_to_column("SampleID")

weighted_unifrac_pcoa <- jaccard$data$Vectors %>%
  dplyr::select(SampleID, PC1, PC2) %>%
  left_join(metadata) %>%
  left_join(shannon)


weighted_unifrac_pcoa <- jaccard$data$Vectors %>%
  dplyr::select(SampleID, PC1, PC2) %>%
  left_join(metadata) %>%
  left_join(shannon) %>%
  ggplot(aes_string(x="PC1", y="PC2", color=ioi))+
  geom_point() +
  geom_convexhull(alpha=0.5, aes_string(fill=ioi)) +
  theme_q2r() +
  scale_color_discrete(name=ioi)

weighted_unifrac_pcoa

ggsave("weighted_unifrac_pcoa_ordination.png", plot=weighted_unifrac_pcoa, path=outdir)

weighted_unifrac_pcoa_no_polygon <- jaccard$data$Vectors %>%
  dplyr::select(SampleID, PC1, PC2) %>%
  left_join(metadata) %>%
  left_join(shannon) %>%
  ggplot(aes_string(x="PC1", y="PC2", color=ioi))+
  geom_point() +
  theme_q2r() +
  scale_color_discrete(name=ioi)

weighted_unifrac_pcoa_no_polygon
ggsave("weighted_unifrac_pcoa_ordination_no_polygons.png", plot=weighted_unifrac_pcoa_no_polygon, path=outdir)
```




### 6.5 Principal Components Analysis

```{r PCA created by ampvis2, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
devtools::source_gist("8d0ca4206a66be7ff6d76fc4ab8e66c6")

cycle_1 <- qza_to_phyloseq(table_dada2,rooted_tree,taxonomy_file,metadata_file) 
cycle_1@sam_data[[ioi]] <- factor(cycle_1@sam_data[[ioi]], levels=ioi_ord)
cycle_1_amp <- phyloseq_to_ampvis2(cycle_1)

euclidean_frame <- amp_ordinate(cycle_1_amp, filter_species = 0, type = "PCA", transform = "hellinger", print_caption = T,sample_color_by = ioi, sample_colorframe = T, detailed_output = T,envfit_factor =ioi,envfit_show = F) 
euclidean_frame$plot
print(euclidean_frame$evf_factor_model)
print(euclidean_frame$screeplot)
ggsave("PCA_ordination_with_boundries.png", plot=euclidean_frame$plot , path=outdir)

loading <- as.data.frame(euclidean_frame$dspecies)
#loading <- select(loading, "PC1", "PC2" , "Genus")
loading <- subset(loading, select = c(PC1, PC2, Genus))
loading <- melt(loading, id.vars="Genus")
loading <- arrange(loading, desc(abs(value)))

loading$Genus <- factor(loading$Genus, levels= unique(loading$Genus))

cutoff <- summary(loading$value)[2]

loading2 <- subset(loading, abs(value) > abs(cutoff))

loading <- loading2[1:100,]

p <- ggplot(data=loading, aes(x=Genus, y=value, fill=variable))+
  geom_bar(stat="identity", width=0.4,position = position_dodge(width = 0.5)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
p
ggsave("Loading_top_100_genera.png", plot=p , path=outdir)


euclidean <- amp_ordinate(cycle_1_amp, filter_species = 0, type = "PCA", transform = "hellinger", print_caption = T,sample_color_by = ioi) 
euclidean

ggsave("PCA_ordination_without_boundries.png", plot=euclidean , path=outdir)
```




## 7 Rarefaction Curves

```{r rarefaction curves, echo=FALSE, fig.height=10, fig.width=16, message=FALSE, warning=FALSE}


numextract <- function(string){
  str_extract(string, "\\-*\\d+\\.*\\d*")
}

obs_rarefy <- read.csv("./qiime/alpha-rareplot/observed_features.csv")
samples <- obs_rarefy[1]
one <- obs_rarefy[2:11]
two <- obs_rarefy[12:21]
three <- obs_rarefy[22:31]
four <- obs_rarefy[32:41]
five <- obs_rarefy[42:51]
six <- obs_rarefy[52:61]
seven <- obs_rarefy[62:71]
eight <- obs_rarefy[72:81]
nine <- obs_rarefy[82:91]
ten <- obs_rarefy[92:101]
obs_item_of_interest <- obs_rarefy[[ioi]]



one_d <- numextract(colnames(obs_rarefy[2:11])[1])
two_d <- numextract(colnames(obs_rarefy[12:21])[1])
three_d <- numextract(colnames(obs_rarefy[22:31])[1])
four_d <- numextract(colnames(obs_rarefy[32:41])[1])
five_d <- numextract(colnames(obs_rarefy[42:51])[1])
six_d <- numextract(colnames(obs_rarefy[52:61])[1])
seven_d <- numextract(colnames(obs_rarefy[62:71])[1])
eight_d <- numextract(colnames(obs_rarefy[72:81])[1])
nine_d <- numextract(colnames(obs_rarefy[82:91])[1])
ten_d <- numextract(colnames(obs_rarefy[92:101])[1])


one["sample"] <- samples
one_flat <- melt(one) %>%
  rename(depth = variable)
one_flat["depth"] <- rep(one_d, dim(one_flat)[1])

one_flat_mean <- one_flat %>%
  group_by(sample) %>%
  summarise(average=mean(value,na.rm = TRUE)) %>%
  ungroup()

one_flat_mean["depth"] <- rep(one_d, dim(one_flat_mean)[1])

two["sample"] <- samples
two_flat <- melt(two) %>%
  rename(depth = variable)
two_flat["depth"] <- rep(two_d, dim(two_flat)[1])

two_flat_mean <- two_flat %>%
  group_by(sample) %>%
  summarise(average=mean(value,na.rm = TRUE)) %>%
  ungroup()

two_flat_mean["depth"] <- rep(two_d, dim(two_flat_mean)[1])

three["sample"] <- samples
three_flat <- melt(three) %>% 
  rename(depth = variable)
three_flat["depth"] <- rep(three_d, dim(three_flat)[1])

three_flat_mean <- three_flat %>%
  group_by(sample) %>%
  summarise(average=mean(value,na.rm = TRUE)) %>%
  ungroup()

three_flat_mean["depth"] <- rep(three_d, dim(three_flat_mean)[1])

four["sample"] <- samples
four_flat <- melt(four) %>%
  rename(depth = variable)
four_flat["depth"] <- rep(four_d, dim(four_flat)[1])

four_flat_mean <- four_flat %>%
  group_by(sample) %>% 
  summarise(average=mean(value,na.rm = TRUE)) %>%
  ungroup()

four_flat_mean["depth"] <- rep(four_d, dim(four_flat_mean)[1])

five["sample"] <- samples
five_flat <- melt(five) %>%
  rename(depth=variable)
five_flat["depth"] <- rep(five_d, dim(five_flat)[1])

five_flat_mean <- five_flat %>%
  group_by(sample) %>%
  summarise(average=mean(value,na.rm = TRUE)) %>%
  ungroup()

five_flat_mean["depth"] <- rep(five_d, dim(five_flat_mean)[1])

six["sample"]<-samples
six_flat <- melt(six)%>%
  rename(depth=variable)
six_flat["depth"]<- rep(six_d, dim(six_flat)[1])

six_flat_mean <- six_flat %>%
  group_by(sample) %>%
  summarise(average=mean(value,na.rm = TRUE)) %>%
  ungroup()

six_flat_mean["depth"] <- rep(six_d, dim(six_flat_mean)[1])

seven["sample"] <- samples
seven_flat <- melt(seven) %>%
  rename(depth=variable)
seven_flat["depth"]<-rep(seven_d, dim(seven_flat)[1])

seven_flat_mean <- seven_flat %>%
  group_by(sample) %>%
  summarise(average=mean(value,na.rm = TRUE)) %>%
  ungroup()

seven_flat_mean["depth"] <- rep(seven_d, dim(seven_flat_mean)[1])

eight["sample"] <- samples
eight_flat <- melt(eight) %>%
  rename(depth=variable)
eight_flat["depth"] <- rep(eight_d, dim(eight_flat)[1])

eight_flat_mean <- eight_flat %>%
  group_by(sample) %>%
  summarise(average=mean(value,na.rm = TRUE)) %>%
  ungroup()

eight_flat_mean["depth"] <- rep(eight_d, dim(eight_flat_mean)[1])

nine["sample"] <- samples
nine_flat <- melt(nine) %>%
  rename(depth=variable)
nine_flat["depth"] <- rep(nine_d, dim(nine_flat)[1])

nine_flat_mean <- nine_flat %>%
  group_by(sample) %>%
  summarise(average=mean(value,na.rm = TRUE)) %>%
  ungroup()

nine_flat_mean["depth"] <- rep(nine_d, dim(nine_flat_mean)[1])

ten["sample"] <- samples
ten_flat <- melt(ten) %>%
  rename(depth=variable)
ten_flat["depth"] <- rep(ten_d, dim(ten_flat)[1])

ten_flat_mean <- ten_flat %>%
  group_by(sample) %>%
  summarise(average=mean(value,na.rm = TRUE)) %>%
  ungroup()

ten_flat_mean["depth"] <- rep(ten_d, dim(ten_flat_mean)[1])


bound <- rbind(one_flat,two_flat, three_flat, four_flat, five_flat, six_flat, seven_flat, eight_flat, nine_flat, ten_flat)
bound_mean <- rbind(one_flat_mean, two_flat_mean, three_flat_mean, four_flat_mean, five_flat_mean, six_flat_mean, seven_flat_mean, eight_flat_mean, nine_flat_mean, ten_flat_mean)



bound$depth <- as.numeric(as.character(bound$depth))

bound <- bound[order(bound$depth),]

rarefy_curve_by_sample <- ggplot(bound, aes(x=depth, y=value, fill=sample, color=sample))+
  geom_point(data=bound_mean, mapping=aes(x=as.numeric(depth), y=as.numeric(average)),size=3) + 
  geom_line(data=bound_mean, mapping= aes(x=as.numeric(depth), y=as.numeric(average), group=sample), size=1) + 
  scale_x_binned("Sampling Depth",n.breaks = 12, nice.breaks = TRUE) + theme_bw() +
  scale_y_continuous("Observed Species") + 
  theme(text=element_text(size = 20))
rarefy_curve_by_sample

ggsave("rarefaction_curve_by_sample.png", plot=rarefy_curve_by_sample, path=outdir)

samples <- obs_rarefy[1]
one <- obs_rarefy[2:11]
two <- obs_rarefy[12:21]
three <- obs_rarefy[22:31]
four <- obs_rarefy[32:41]
five <- obs_rarefy[42:51]
six <- obs_rarefy[52:61]
seven <- obs_rarefy[62:71]
eight <- obs_rarefy[72:81]
nine <- obs_rarefy[82:91]
ten <- obs_rarefy[92:101]
obs_item_of_interest <- obs_rarefy[[ioi]]



one_d <- numextract(colnames(obs_rarefy[2:11])[1])
two_d <- numextract(colnames(obs_rarefy[12:21])[1])
three_d <- numextract(colnames(obs_rarefy[22:31])[1])
four_d <- numextract(colnames(obs_rarefy[32:41])[1])
five_d <- numextract(colnames(obs_rarefy[42:51])[1])
six_d <- numextract(colnames(obs_rarefy[52:61])[1])
seven_d <- numextract(colnames(obs_rarefy[62:71])[1])
eight_d <- numextract(colnames(obs_rarefy[72:81])[1])
nine_d <- numextract(colnames(obs_rarefy[82:91])[1])
ten_d <- numextract(colnames(obs_rarefy[92:101])[1])

fct_ioi <- factor(obs_item_of_interest, levels=ioi_ord)
one <- cbind(one,fct_ioi)
one_flat <- melt(one,id.vars = "fct_ioi") %>%
  rename(depth = variable) %>%
  rename(ioi = fct_ioi)
one_flat["depth"] <- rep(one_d, dim(one_flat)[1])

one_flat_mean <- one_flat %>%
  group_by(ioi) %>%
  summarise(average=mean(value,na.rm = TRUE)) %>%
  ungroup()

one_flat_mean["depth"] <- rep(one_d, dim(one_flat_mean)[1])

two <- cbind(two, fct_ioi)
two_flat <- melt(two, id.vars = "fct_ioi") %>%
  rename(depth = variable) %>%
  rename(ioi = fct_ioi)
two_flat["depth"] <- rep(two_d, dim(two_flat)[1])

two_flat_mean <- two_flat %>%
  group_by(ioi) %>%
  summarise(average=mean(value,na.rm = TRUE)) %>%
  ungroup()

two_flat_mean["depth"] <- rep(two_d, dim(two_flat_mean)[1])

three <- cbind(three, fct_ioi)
three_flat <- melt(three, id.vars = "fct_ioi") %>% 
  rename(depth = variable) %>%
  rename(ioi = fct_ioi)
three_flat["depth"] <- rep(three_d, dim(three_flat)[1])

three_flat_mean <- three_flat %>%
  group_by(ioi) %>%
  summarise(average=mean(value,na.rm = TRUE)) %>%
  ungroup()

three_flat_mean["depth"] <- rep(three_d, dim(three_flat_mean)[1])

four<- cbind(four, fct_ioi)
four_flat <- melt(four, id.vars = "fct_ioi") %>%
  rename(depth = variable) %>%
  rename(ioi = fct_ioi)
four_flat["depth"] <- rep(four_d, dim(four_flat)[1])

four_flat_mean <- four_flat %>%
  group_by(ioi) %>% 
  summarise(average=mean(value,na.rm = TRUE)) %>%
  ungroup()

four_flat_mean["depth"] <- rep(four_d, dim(four_flat_mean)[1])

five <- cbind(five, fct_ioi)
five_flat <- melt(five, id.vars = "fct_ioi") %>%
  rename(depth=variable) %>%
  rename(ioi= fct_ioi)
five_flat["depth"] <- rep(five_d, dim(five_flat)[1])

five_flat_mean <- five_flat %>%
  group_by(ioi) %>%
  summarise(average=mean(value,na.rm = TRUE)) %>%
  ungroup()

five_flat_mean["depth"] <- rep(five_d, dim(five_flat_mean)[1])

six<-cbind(six, fct_ioi)
six_flat <- melt(six, id.vars = "fct_ioi")%>%
  rename(depth=variable) %>%
  rename(ioi = fct_ioi)
six_flat["depth"]<- rep(six_d, dim(six_flat)[1])

six_flat_mean <- six_flat %>%
  group_by(ioi) %>%
  summarise(average=mean(value,na.rm = TRUE)) %>%
  ungroup()

six_flat_mean["depth"] <- rep(six_d, dim(six_flat_mean)[1])

seven<- cbind(seven, fct_ioi)
seven_flat <- melt(seven,id.vars = "fct_ioi") %>%
  rename(depth=variable) %>%
  rename(ioi = fct_ioi)
seven_flat["depth"]<-rep(seven_d, dim(seven_flat)[1])

seven_flat_mean <- seven_flat %>%
  group_by(ioi) %>%
  summarise(average=mean(value,na.rm = TRUE)) %>%
  ungroup()

seven_flat_mean["depth"] <- rep(seven_d, dim(seven_flat_mean)[1])

eight<- cbind(eight, fct_ioi)
eight_flat <- melt(eight,id.vars = "fct_ioi") %>%
  rename(depth=variable) %>%
  rename(ioi = fct_ioi)
eight_flat["depth"] <- rep(eight_d, dim(eight_flat)[1])

eight_flat_mean <- eight_flat %>%
  group_by(ioi) %>%
  summarise(average=mean(value,na.rm = TRUE)) %>%
  ungroup()

eight_flat_mean["depth"] <- rep(eight_d, dim(eight_flat_mean)[1])

nine <- cbind(nine, fct_ioi)
nine_flat <- melt(nine,id.vars = "fct_ioi") %>%
  rename(depth=variable) %>%
  rename(ioi = fct_ioi)
nine_flat["depth"] <- rep(nine_d, dim(nine_flat)[1])

nine_flat_mean <- nine_flat %>%
  group_by(ioi) %>%
  summarise(average=mean(value,na.rm = TRUE)) %>%
  ungroup()

nine_flat_mean["depth"] <- rep(nine_d, dim(nine_flat_mean)[1])

ten <- cbind(ten,fct_ioi)
ten_flat <- melt(ten, id.vars = "fct_ioi") %>%
  rename(depth=variable) %>%
  rename(ioi = fct_ioi)
ten_flat["depth"] <- rep(ten_d, dim(ten_flat)[1])

ten_flat_mean <- ten_flat %>%
  group_by(ioi) %>%
  summarise(average=mean(value,na.rm = TRUE)) %>%
  ungroup()

ten_flat_mean["depth"] <- rep(ten_d, dim(ten_flat_mean)[1])


bound <- rbind(one_flat,two_flat, three_flat, four_flat, five_flat, six_flat, seven_flat, eight_flat, nine_flat, ten_flat)
bound_mean <- rbind(one_flat_mean, two_flat_mean, three_flat_mean, four_flat_mean, five_flat_mean, six_flat_mean, seven_flat_mean, eight_flat_mean, nine_flat_mean, ten_flat_mean)

bound$depth <- as.numeric(as.character(bound$depth))

bound <- bound[order(bound$depth),]


rarefy_curve_by_treatment <- ggplot(bound, aes(x=depth, y=value, color=fct_relevel(fct_reorder(ioi, as.numeric(ioi)))))+
  geom_point(data=bound_mean, mapping=aes(x=as.numeric(depth), y=as.numeric(average)),size=3) + 
  geom_line(data=bound_mean, mapping= aes(x=as.numeric(depth), y=as.numeric(average), group=ioi), size=1) + 
  scale_x_binned("Sampling Depth",n.breaks = 12, nice.breaks = TRUE) + theme_bw() +
  scale_y_continuous("Observed Species") + 
  theme(text=element_text(size = 20)) + 
  scale_color_discrete(name = ioi)

rarefy_curve_by_treatment
ggsave("rarefaction_curve_by_treatment.png", plot=rarefy_curve_by_treatment, path=outdir)
```




## 8 Ranked Abundance Curves

```{r, echo=FALSE, fig.height=10, fig.width=16, message=FALSE, warning=FALSE}
library(ampvis2)
library(ggplot2)
devtools::source_gist("8d0ca4206a66be7ff6d76fc4ab8e66c6")

cycle_1 <- qza_to_phyloseq(table_dada2,rooted_tree,taxonomy_file,metadata_file) 
cycle_1_amp <- phyloseq_to_ampvis2(cycle_1)

by_sample <- ampvis2::amp_rankabundance(cycle_1_amp, group_by = "SampleID") + theme(text=element_text(size = 20))
by_sample 
ggsave("rank_abundance_curve_by_sample.png", plot=by_sample, path=outdir)

cycle_1_amp$metadata[[ioi]] <- factor(cycle_1_amp$metadata[[ioi]], levels=ioi_ord)

by_group <- ampvis2::amp_rankabundance(cycle_1_amp, group_by = ioi) + theme(text=element_text(size = 20))
by_group
ggsave("rank_abundance_curve_by_treatemnt.png", plot=by_group, path=outdir)




```




## 9 Unifrac Heatmap 

```{r echo=FALSE,fig.height=8, fig.width=14, message=FALSE, warning=FALSE}

metadata<-read_q2metadata(metadata_file)
#metadata[[ioi]] <- fct_reorder(metadata[[ioi]])
w_unifrac_dist <- qiime2R::read_qza("./qiime/core-metric-results/weighted_unifrac_distance_matrix.qza")
u_unifrac_dist <- qiime2R::read_qza("./qiime/core-metric-results/unweighted_unifrac_distance_matrix.qza")

un_weight_matrix <- as.matrix(u_unifrac_dist$data)
un_weight_matrix[lower.tri(un_weight_matrix)] <- NA
un_weight_heatmap <- pheatmap(un_weight_matrix, cluster_rows = F, cluster_cols = F, na_col = "grey", display_numbers = T, number_format ="%.3f",main = "Beta Diversity Heatmap using Unweighted Unifrac Distance")
un_weight_heatmap

ggsave("beta_div_heatmap_unweight_unifrac.png", plot=un_weight_heatmap, path=outdir)

weight_matrix <- as.matrix(w_unifrac_dist$data)
weight_matrix[lower.tri(weight_matrix)] <- NA
weight_heatmap <- pheatmap(weight_matrix, cluster_rows = F, cluster_cols = F, na_col = "grey", display_numbers = T, number_format ="%.3f",main = "Beta Diversity Heatmap using Weighted Unifrac Distance")
weight_heatmap

ggsave("beta_div_heatmap_Weighted_unifrac.png", plot=weight_heatmap, path=outdir)
```




## 10 Beta Diversity Between Groups

### 10.1 Weighted boxplots

```{r generated Weighted boxplots, echo=FALSE, message=FALSE, warning=FALSE}
weighted_measurements <- read.csv("./qiime/weighted-sig/raw_data.tsv", sep="\t")

split_weighted_measure <- split(weighted_measurements, weighted_measurements$Group1)

for(frame in split_weighted_measure){
  
  frame$Group2 <- factor(frame$Group2, levels=ioi_ord)
  p <- ggplot(frame, aes(x=frame$Group2, y=Distance)) + geom_violin()+ geom_boxplot(width=0.1)
  p<- p + stat_summary(fun.y=mean, geom="point", shape=23, size=2)
  p<-p + stat_summary(fun.y=median, geom="point", size=2, color="red")
  p <- p+geom_jitter(shape=16, position=position_jitter(0.2)) + xlab(ioi) + ylab("Weighted Unifrac Distance") + theme_bw() + labs(title = (as.character(paste0("Distance to ",frame$Group1[1]))))
  print(p)
  ggsave(paste0("weighted_unifrac_boxplot_",frame$Group1[1],".png"), plot=p,path=outdir)
}
```




### 10.2 Unweighted boxplots

```{r generated Unweighted boxplots, echo=FALSE, message=FALSE, warning=FALSE}
unweighted_measurements <- read.csv("./qiime/unweighted-sig/raw_data.tsv", sep="\t")

split_data_frames <- split(unweighted_measurements, unweighted_measurements$Group1)


for(frame in split_data_frames){
  
  frame$Group2 <- factor(frame$Group2, levels=ioi_ord)
  p <- ggplot(frame, aes(x=frame$Group2, y=Distance)) + geom_violin()+ geom_boxplot(width=0.1)
  p<- p + stat_summary(fun.y=mean, geom="point", shape=23, size=2)
  p<-p + stat_summary(fun.y=median, geom="point", size=2, color="red")
  p <- p+geom_jitter(shape=16, position=position_jitter(0.2)) + xlab(ioi) + ylab("Unweighted Unifrac Distance") + theme_bw() + labs(title = (as.character(paste0("Distance to ",frame$Group1[1]))))
  print(p)
  ggsave(paste0("unweighted_unifrac_boxplot_",frame$Group1[1],".png"), plot=p,path=outdir)
}
```




## 11 UPGMA Plots
### 11.1 UPGMA Plot for Weighted Unifrac Distance

```{r UPGMA Plot of Weighted Unifrac Distance, echo=FALSE, fig.height=8, fig.width=10, message=FALSE, warning=FALSE}
library("ggplot2")
library("ggdendro")
cycle_1 <- qza_to_phyloseq(table_dada2,rooted_tree,taxonomy_file,metadata_file)
w_unifrac <- UniFrac(cycle_1,weighted=TRUE)


hclust_res <- hclust(w_unifrac, method="average")

hclust_dend <- as.dendrogram(hclust_res)
ddata <- dendro_data(hclust_dend, type="rectangle")
p <- ggplot(segment(ddata)) +
  geom_segment(aes(x=x,y=y, xend=xend, yend=yend)) +
  coord_flip() +
  scale_y_reverse(expand=c(0.2,0)) + 
  geom_text(data=ddata$labels,aes(x=x,y=y,label=label, hjust=0)) + 
  theme(text=element_text(size = 12, face = "plain")) +
  theme_minimal() + 
  theme(axis.title.y = element_blank(), 
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  ylab("Weighted Unifrac Distance") 
  
#sample_data(cycle_1)[[ioi]] <- reorder(sample_data(cycle_1)[[ioi]])

level <- "Phylum"
#Select number of taxa at phylum level to plot
topN <- 10

#selects top n taxa and creates phyloseq object
cycle_n_names <- sort(tapply(taxa_sums(cycle_1), tax_table(cycle_1)[,as.character(level)], sum), TRUE) [1:topN]
cycle_n <- subset_taxa(cycle_1, get(level) %in% names(cycle_n_names))
cycle_n_glom <- tax_glom(cycle_n, taxrank = level)

#selects top n + 1 to the end of table
rest_tax <- sort(tapply(taxa_sums(cycle_1), tax_table(cycle_1)[,as.character(level)], sum), TRUE) [topN+1:length(tax_table(cycle_1))]
rest <- subset_taxa(cycle_1, get(level) %in% names(rest_tax))
rest_glom <- tax_glom(rest, taxrank = level)

#collapse all phyla not in top 10 into one taxon
rest_merge <- merge_taxa(rest, taxa_names(rest))

#create dummy taxonomy classification for collapsed taxa
taxvec <- c("Other","Other","Other","Other","Other", "Other","Other")
taxlist <- list(OTU1=parse_taxonomy_default(taxvec))
tax_tab <- build_tax_table(taxlist)
colnames(tax_tab) <- c("Kingdom","Phylum","Class","Order","Family","Genus","Species")
rownames(tax_tab) <- rownames(tax_table(rest_merge))
tax_table(rest_merge) <- tax_tab

#attaches top n phyloseq object with the other phyloseq object
merge_tax <- merge_phyloseq(tax_table(cycle_n), tax_table(rest_merge))
merge_otu <- merge_phyloseq(otu_table(cycle_n), otu_table(rest_merge))
merge_samp <- merge_phyloseq(sample_data(cycle_n), sample_data(rest_merge))

merged <- merge_phyloseq(merge_tax, merge_otu, merge_samp)

#calculates relative abundance based on sample provided and turns that into a dataframe for ggplot2 plotting
cycle_1_phylum <- merged %>%
  tax_glom(taxrank = level) %>%
  transform_sample_counts(function(x) {x/sum(x)}) %>%
  psmelt() %>%
  arrange(Abundance) 

cycle_1_phylum$Sample <- as.character(cycle_1_phylum$Sample)
cycle_1_phylum$Sample <- factor(cycle_1_phylum$Sample, levels=ddata$labels$label)


#ggplot2 stacked bar chart separated by sample in the metadata from qiime2
by_sample <- ggplot(cycle_1_phylum, aes(x=Sample, y=Abundance, fill=fct_relevel(fct_reorder(cycle_1_phylum[[level]], Abundance),"Other"))) + 
  geom_bar(stat= "identity") +
  guides(fill=guide_legend(title=level))+
  theme_minimal() +
  xlab(element_blank())+
  theme(axis.text.x = element_text(vjust=0.5, hjust=1))+
  coord_flip() +
  theme(text=element_text(size = 12,face = "plain"),axis.title.y = element_text(color="Black"), axis.text.y = element_text(color="Black")) + 
  ylab("Relative Abundance at Phylum Level")
  


arr <- ggarrange(p, by_sample, nrow=1)
weight_plot <- annotate_figure(arr,  top="UPGMA cluster tree based on Weighted Unifrac distance")
weight_plot

ggsave("UPGMA_weighted_unifrac_plot.png", plot=weight_plot, path=outdir)
```




### 11.2 UPGMA Plot for Unweighted Unifrac Distance

```{r UPGMA Plot of Unweighted Unifrac Distances, echo=FALSE, fig.height=8, fig.width=10, message=FALSE, warning=FALSE}
cycle_1 <- qza_to_phyloseq(table_dada2,rooted_tree,taxonomy_file,metadata_file)
u_unifrac <- UniFrac(cycle_1, weighted = FALSE)


hclust_res <- hclust(u_unifrac, method="average")

hclust_dend <- as.dendrogram(hclust_res)
ddata <- dendro_data(hclust_dend, type="rectangle")
p <- ggplot(segment(ddata)) +
  geom_segment(aes(x=x,y=y, xend=xend, yend=yend)) +
  coord_flip() +
  scale_y_reverse(expand=c(0.2,0)) + 
  geom_text(data=ddata$labels,aes(x=x,y=y,label=label, hjust=0)) + 
  theme(text=element_text(size = 12, face = "plain")) +
  theme_minimal() + 
  theme(axis.title.y = element_blank(), 
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  ylab("Unweighted Unifrac Distance") 
  
#sample_data(cycle_1)[[ioi]] <- reorder(sample_data(cycle_1)[[ioi]])

level <- "Phylum"
#Select number of taxa at phylum level to plot
topN <- 10

#selects top n taxa and creates phyloseq object
cycle_n_names <- sort(tapply(taxa_sums(cycle_1), tax_table(cycle_1)[,as.character(level)], sum), TRUE) [1:topN]
cycle_n <- subset_taxa(cycle_1, get(level) %in% names(cycle_n_names))
cycle_n_glom <- tax_glom(cycle_n, taxrank = level)

#selects top n + 1 to the end of table
rest_tax <- sort(tapply(taxa_sums(cycle_1), tax_table(cycle_1)[,as.character(level)], sum), TRUE) [topN+1:length(tax_table(cycle_1))]
rest <- subset_taxa(cycle_1, get(level) %in% names(rest_tax))
rest_glom <- tax_glom(rest, taxrank = level)

#collapse all phyla not in top 10 into one taxon
rest_merge <- merge_taxa(rest, taxa_names(rest))

#create dummy taxonomy classification for collapsed taxa
taxvec <- c("Other","Other","Other","Other","Other", "Other","Other")
taxlist <- list(OTU1=parse_taxonomy_default(taxvec))
tax_tab <- build_tax_table(taxlist)
colnames(tax_tab) <- c("Kingdom","Phylum","Class","Order","Family","Genus","Species")
rownames(tax_tab) <- rownames(tax_table(rest_merge))
tax_table(rest_merge) <- tax_tab

#attaches top n phyloseq object with the other phyloseq object
merge_tax <- merge_phyloseq(tax_table(cycle_n), tax_table(rest_merge))
merge_otu <- merge_phyloseq(otu_table(cycle_n), otu_table(rest_merge))
merge_samp <- merge_phyloseq(sample_data(cycle_n), sample_data(rest_merge))

merged <- merge_phyloseq(merge_tax, merge_otu, merge_samp)

#calculates relative abundance based on sample provided and turns that into a dataframe for ggplot2 plotting
cycle_1_phylum <- merged %>%
  tax_glom(taxrank = level) %>%
  transform_sample_counts(function(x) {x/sum(x)}) %>%
  psmelt() %>%
  arrange(Abundance) 

cycle_1_phylum$Sample <- as.character(cycle_1_phylum$Sample)
cycle_1_phylum$Sample <- factor(cycle_1_phylum$Sample, levels=ddata$labels$label)

#ggplot2 stacked bar chart separated by sample in the metadata from qiime2
by_sample <- ggplot(cycle_1_phylum, aes(x=Sample, y=Abundance, fill=fct_relevel(fct_reorder(cycle_1_phylum[[level]], Abundance),"Other"))) + 
  geom_bar(stat= "identity") +
  guides(fill=guide_legend(title=level))+
  theme_minimal() +
  xlab(element_blank())+
  theme(axis.text.x = element_text(vjust=0.5, hjust=1))+
  coord_flip() +
  theme(text=element_text(size = 12, face = "plain"),axis.title.x = element_text(color="Black"), axis.text.y = element_text(color="Black")) + 
  ylab("Relative Abundance at Phylum Level")
  


arr <- ggarrange(p, by_sample, nrow=1)
unweight_plot <- annotate_figure(arr,  top="UPGMA cluster tree based on Unweighted Unifrac distance")
unweight_plot

ggsave("UPGMA_unweighted_unifrac_plot.png", plot=unweight_plot, path=outdir)

```




## 12 Beta Diversity Statistics

### 12.1 PERMANOVA Weighted Unifrac

```{r Methods to do pairwise PERMANOVA comparisons, message=FALSE, warning=FALSE, include=FALSE}
pairwise_permanova <- function(dat, column_of_int, dissim){
  all_comparisons <- unique(sample_data(dat)[[column_of_int]])
  combos <- (t(combn(all_comparisons,2)))
  p_res <- c()
  
  for(i in 1:dim(combos)[1]){
    comp <- prune_samples(sample_data(dat)[[ioi]] %in% combos[i,], dat)
    grp <- get_variable(comp, column_of_int)
  
    perma <- PERMANOVA::PerMANOVA.Simple(D = as.matrix(distance(comp, dissim)) , grupo = grp)
    permanova <- tibble::tibble(data.frame(perma$Inicial))
    permanova_stats <- data.frame(as.character(combos[i,1]), as.character(combos[i,2]), permanova[1,], perma$pval)
    colnames(permanova_stats)[1] <- "group1"
    colnames(permanova_stats)[2] <- "group2"
    colnames(permanova_stats)[10] <- "p"
    data.frame(permanova_stats)
    p_res <- rbind(p_res, permanova_stats)
  
  }
  tmp_2 <- data.frame(p_res)
  #colnames(tmp_2)<- c("group1","group2", "statistic", "permutations", "p")
  result <- adjust_pvalue(tmp_2, p.col = "p", output.col = "p.adj",method = "BH")
  result <- add_significance(result, p.col="p.adj",output.col = "p.signif",symbols =c("****", "***", "**", "$*$", "ns"))
  res_df <- data.frame(result)
  res_df <- res_df[mixedorder(res_df$group1),]
  return(res_df)
  
}

#out <- pairwise_anosim(cycle_1, ioi , dissim = "bray")
#out
```


```{r Methods to do pairwise Unifrac PERMANOVA comparisons, message=FALSE, warning=FALSE, include=FALSE}
library(phyloseq)
pairwise_permanova_uni <- function(dat, column_of_int, weight){
  all_comparisons <- unique(sample_data(dat)[[column_of_int]])
  
  comb_tab <- (t(combn(all_comparisons,2)))
  
  p_res <- c()


  for(i in 1:dim(comb_tab)[1]){
    try <- sample_data(dat)[[ioi]]
    #comp1 <- subset_samples(dat, eval(as.name(ioi)) %in% comb_tab[i,])
    comp <- prune_samples(sample_data(dat)[[ioi]] %in% comb_tab[i,], dat)
    grp <- get_variable(comp, column_of_int)
  
    tmp <- UniFrac(comp, weighted = weight)
    perma <- PERMANOVA::PerMANOVA.Simple(D = as.matrix(tmp), grupo = grp)
    permanova <- tibble::tibble(data.frame(perma$Inicial))
    permanova_stats <- data.frame(as.character(comb_tab[i,1]), as.character(comb_tab[i,2]), permanova[1,], perma$pval)
    colnames(permanova_stats)[1] <- "group1"
    colnames(permanova_stats)[2] <- "group2"
    colnames(permanova_stats)[10] <- "p"
    data.frame(permanova_stats)
    p_res <- rbind(p_res, permanova_stats)
  
  }
  tmp_2 <- data.frame(p_res)
  #colnames(tmp_2)<- c("group1","group2", "statistic", "permutations", "p")
  result <- adjust_pvalue(tmp_2, p.col = "p", output.col = "p.adj",method = "BH")
  result <- add_significance(result, p.col="p.adj",output.col = "p.signif",symbols =c("****", "***", "**", "$*$", "ns"))
  res_df <- data.frame(result)
  res_df <- res_df[mixedorder(res_df$group1),]
  return(res_df)
  
}

```


```{r Weighted Unifrac PERMANOVA, message=FALSE, warning=FALSE, include=FALSE, results='markup'}
library(PERMANOVA)

cycle_1 <- qza_to_phyloseq(table_dada2,rooted_tree,taxonomy_file,metadata_file)
sample_data(cycle_1)[[ioi]] <- factor(sample_data(cycle_1)[[ioi]], levels=ioi_ord)

# PERMANOVA of Weighted Unifrac
u_unifrac <- UniFrac(cycle_1, weighted = T,)

res <- PERMANOVA::PerMANOVA.Simple(D = as.matrix(u_unifrac), grupo = cycle_1@sam_data[[ioi]])

perma_res <- data.frame(data.frame(res$Inicial),res$pval)

# Pairwise PERMANOVA of Weighted Unifrac
test <- pairwise_permanova_uni(cycle_1, ioi, T)

perma_pairwise <- kable(test) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))


```


```{r print weighted unifrac PERMANOVA, echo=FALSE, message=FALSE, warning=FALSE}
kable(perma_res) %>%
  kable_styling(bootstrap_options = c( "hover", "condensed", "responsive"))
  

perma_pairwise

```

### 12.2 PERMANOVA Unweighted Unifrac 

```{r Permanova results Weighted Unifrac from qiime, message=FALSE, warning=FALSE, include=FALSE}
cycle_1 <- qza_to_phyloseq(table_dada2,rooted_tree,taxonomy_file,metadata_file)
sample_data(cycle_1)[[ioi]] <- factor(sample_data(cycle_1)[[ioi]], levels=ioi_ord)
# PERMANOVA of Unweighted Unifrac
u_unifrac <- UniFrac(cycle_1, weighted = F)
grp <- get_variable(cycle_1, ioi)

res <- PERMANOVA::PerMANOVA.Simple(D = as.matrix(u_unifrac), grupo = cycle_1@sam_data[[ioi]])

perma_res <- data.frame(data.frame(res$Inicial),res$pval)

# Pairwise PERMANOVA of Unweighted Unifrac
test <- pairwise_permanova_uni(cycle_1, ioi, F)
perma_pairwise <- kable(test) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

```{r print unweighted unifrac PERMANOVA, echo=FALSE, message=FALSE, warning=FALSE}
kable(perma_res) %>%
  kable_styling(bootstrap_options = c( "hover", "condensed", "responsive"))

perma_pairwise

```



### 12.3 PERMANOVA Bray Curtis

```{r Bray Curtis PERMANOVA, message=FALSE, warning=FALSE, include=FALSE}
cycle_1 <- qza_to_phyloseq(table_dada2,rooted_tree,taxonomy_file,metadata_file)
sample_data(cycle_1)[[ioi]] <- factor(sample_data(cycle_1)[[ioi]], levels=ioi_ord)
#PERMANOVA Bray Curtis
grp <- get_variable(cycle_1, ioi)

res <- PERMANOVA::PerMANOVA.Simple(D = as.matrix(distance(cycle_1, 'bray',)) , grupo = cycle_1@sam_data[[ioi]])
perma_res <- data.frame(data.frame(res$Inicial),res$pval)

#PERMANOVA Bray Curtis Pairwise
bray_pairwise <- pairwise_permanova(cycle_1,ioi,"bray")
pairwise_perma <- kable(bray_pairwise) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

```{r print bray PERMANOVA, echo=FALSE, message=FALSE, warning=FALSE}
kable(perma_res) %>%
  kable_styling(bootstrap_options = c( "hover", "condensed", "responsive"))

pairwise_perma

```


### 12.4 PERMANOVA Jaccard 

```{r Jaccard PERMANOVA, message=FALSE, warning=FALSE, include=FALSE}
cycle_1 <- qza_to_phyloseq(table_dada2,rooted_tree,taxonomy_file,metadata_file)
sample_data(cycle_1)[[ioi]] <- factor(sample_data(cycle_1)[[ioi]], levels=ioi_ord)
# PERMANOVA Jaccard distance
grp <- get_variable(cycle_1, ioi)

res <- PERMANOVA::PerMANOVA.Simple(D = as.matrix(distance(cycle_1, 'jaccard',)) , grupo = cycle_1@sam_data[[ioi]])
perma_res <- data.frame(data.frame(res$Inicial),res$pval)
# PERMANOVA Pairwise Jaccard 
jaccard_pairwise <- pairwise_permanova(cycle_1, ioi, "jaccard")
pairwise_perma <- kable(jaccard_pairwise) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

```{r print jaccard PERMANOVA, echo=FALSE, message=FALSE, warning=FALSE}
kable(perma_res) %>%
  kable_styling(bootstrap_options = c( "hover", "condensed", "responsive"))

pairwise_perma


```



### 12.5 ANOSIM

```{r Method to do pairwise anosim comparisons, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE, eval=FALSE}
library(rstatix)
library(ggpubr)
pairwise_anosim <- function(dat, column_of_int, dissim){
  all_comparisons <- unique(sample_data(dat)[[column_of_int]])
  combos <- (t(combn(all_comparisons,2)))
  p_res <- c()
  
  for(i in 1:dim(combos)[1]){
    comp <- prune_samples(sample_data(dat)[[ioi]] %in% combos[i,], dat)
    grp <- get_variable(comp, column_of_int)
  
    ano <- anosim(distance(comp, dissim), grp)
    ano_stats <- c(as.character(combos[i,1]), as.character(combos[i,2]), round(ano$statistic,6), ano$permutations, ano$dissimilarity, round(ano$signif,3))
    #tmp <- c(ano, as.character(item))
    p_res <- rbind(p_res, ano_stats)
  
  }
  tmp_2 <- data.frame(p_res)
  colnames(tmp_2)<- c("group1", "group2", "statistic", "permutations", "dissimilarity", "p.val")
  result <- adjust_pvalue(tmp_2, p.col = "p.val", output.col = "p.adjust",method = "BH")
  result <- add_significance(result, p.col = "p.adjust", output.col = "p.signif",symbols =c('****', '***', '**', '$*$', 'ns'))
  result <- add_significance(result, p.col="p.adjust",output.col = "p.signif.box",symbols =c('****', '***', '**', '*', 'ns'))
  res_df <- data.frame(result)
  res_df <- res_df[mixedorder(res_df$group1),]
  return(res_df)
  
}

#out <- pairwise_anosim(cycle_1, ioi , dissim = "bray")
#out

pairwise_anosim_uni <- function(dat, column_of_int, weight){
  all_comparisons <- unique(sample_data(dat)[[column_of_int]])
  combos <- (t(combn(all_comparisons,2)))
  p_res <- c()
  
  for(i in 1:dim(combos)[1]){
    comp <- prune_samples(sample_data(dat)[[ioi]] %in% combos[i,], dat)
    grp <- get_variable(comp, column_of_int)
  
    tmp <- UniFrac(comp, weighted = weight)
    ano <- anosim(tmp, sample_data(comp)[[ioi]])
    ano_stats <- c(as.character(combos[i,1]), as.character(combos[i,2]), round(ano$statistic,6), ano$permutations, ano$dissimilarity, round(ano$signif,3))
    #tmp <- c(ano, as.character(item))
    p_res <- rbind(p_res, ano_stats)
  
  }
  tmp_2 <- data.frame(p_res)
  colnames(tmp_2)<- c("group1","group2", "statistic", "permutations", "p")
  result <- adjust_pvalue(tmp_2, p.col = "p", output.col = "p.adj",method = "BH")
  result <- add_significance(result, p.col="p.adj",output.col = "p.signif",symbols =c('****', '***', '**', '$*$', 'ns'))
  result <- add_significance(result, p.col="p.adj",output.col = "p.signif.box",symbols =c('****', '***', '**', '*', 'ns'))
  res_df <- data.frame(result)
  res_df <- res_df[mixedorder(res_df$group1),]
  return(res_df)
  
}

```


```{r echo=FALSE, fig.height=8, fig.width=10, message=FALSE, warning=FALSE, include=FALSE,eval=FALSE}
library(rstatix)
library(ggpubr)
cycle_1 <- qza_to_phyloseq(table_dada2,rooted_tree,taxonomy_file,metadata_file)
sample_data(cycle_1)[[ioi]] <- factor(sample_data(cycle_1)[[ioi]], levels=ioi_ord)

#Weighted Unifrac Anosim
tmp <- UniFrac(cycle_1, weighted=T)
res <- anosim(tmp, sample_data(cycle_1)[[ioi]])
res


#Weighted Unifrac Anosim Pairwise
w_uni_pear <- pairwise_anosim_uni(cycle_1, ioi, T) 
w_uni_pear %>%
  dplyr::select(group1,group2,statistic,permutations,p,p.adj,p.signif) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

#Weighted Unifrac ANOSIM Plot
#w_uni_plot <- autoplot(res, notch=F) + theme_bw()+labs(title = "ANOSIM Weighted Unifrac") + xlab(ioi) + ylab("Rank Value")  + stat_pvalue_manual(w_uni_pear, label = "p.signif.box", y.position = max(res$dis.rank)-(.5*median(res$dis.rank)),step.increase = .03, tip.length = 0.01, guide="none", hide.ns=T) 
#w_uni_plot
#Save image to disk
#ggsave("ANOSIM_weighted_unifrac_boxplot.png", plot = w_uni_plot,path = outdir)

#Weighted Unifrac ANOSIM Plot No Bracket
w_uni_plot_no_brac <- autoplot(res, notch=F) + theme_bw()+ aes(fill=res$class.vec) +labs(title = "ANOSIM Weighted Unifrac", fill=ioi)+ xlab(as.character(ioi)) + ylab("Rank Value")
w_uni_plot_no_brac
#Save image to disk
ggsave("ANOSIM_weighted_unifrac_boxplot_no_signif.png", plot = w_uni_plot_no_brac,path = outdir)

#Unweighted Unifrac
tmp <- UniFrac(cycle_1, weighted = F)
res <- anosim(tmp, sample_data(cycle_1)[[ioi]])
res
#Unweighted Unifrac Anosim Pairwise
uw_uni_pair <- pairwise_anosim_uni(cycle_1, ioi, T)

uw_uni_pair %>%
  dplyr::select(group1,group2,statistic,permutations,p,p.adj,p.signif) %>%
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

#Unweighted Unifrac ANOSIM plot
#uw_uni_plot <- autoplot(res,notch = F) +theme_bw() + labs(title = "ANOSIM Unweighted Unifrac") +
  stat_pvalue_manual(uw_uni_pair, label="p.signif.box", y.position = max(res$dis.rank)-(.5*median(res$dis.rank)),step.increase = .03, tip.length = 0.01, hide.ns=T)+ xlab(as.character(ioi)) + ylab("Rank Value")
#uw_uni_plot
#save to disk
#ggsave("ANOSIM_unweighted_unifrac_boxplot.png", plot=uw_uni_plot, path=outdir)
#Unweighted Unifrac ANOSIM plot No bracket
uw_uni_plot_no_brac <- autoplot(res,notch = F) +theme_bw() +aes(fill=res$class.vec)+ labs(title = "ANOSIM Unweighted Unifrac", fill=ioi) + xlab(as.character(ioi)) + ylab("Rank Value")
uw_uni_plot_no_brac
#save to disk
ggsave("ANOSIM_unweighted_unifrac_boxplot_no_brackets.png", plot=uw_uni_plot_no_brac, path=outdir)

#Bray Curtis Distance
grp <- get_variable(cycle_1, ioi)
ano <- anosim(distance(cycle_1, "bray"), grp)
ano
#Bray Curtis ANOSIM pairwise
bray_pair <- pairwise_anosim(cycle_1, ioi, "bray")
bray_pair %>%
  dplyr::select(group1,group2,statistic,permutations,dissimilarity,p.val,p.adjust,p.signif) %>%
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
#Bray Curtis ANOSIM plot
#bray_plot <- autoplot(ano, notch=F) + theme_bw() + labs(title ="ANOSIM Bray Curtis") + stat_pvalue_manual(bray_pair, label="p.signif.box", y.position = max(ano$dis.rank)-(.5*median(res$dis.rank)), step.increase = .03, tip.length = 0.01, hide.ns=T) + xlab(as.character(ioi)) + ylab("Rank Value")
#bray_plot
#save to disk
#ggsave("ANOSIM_bray_curtis_boxplot.png", plot=bray_plot, path=outdir)
#Bray Curtis ANOSIM plot No Bracket
bray_plot_no_brac <- autoplot(ano, notch=F) + theme_bw() +aes(fill=ano$class.vec) + labs(title ="ANOSIM Bray Curtis", fill=ioi) + xlab(as.character(ioi)) + ylab("Rank Value")
bray_plot_no_brac
#save to disk
ggsave("ANOSIM_bray_curtis_boxplot_no_bracket.png", plot=bray_plot_no_brac, path=outdir)

#Jaccard Distance
grp <- get_variable(cycle_1, ioi)
ano <- anosim(distance(cycle_1, "jaccard"), grp)
ano
#Jaccard ANOSIM pairwise
jaccard_pair <- pairwise_anosim(cycle_1, ioi, "jaccard")
jaccard_pair %>%
  dplyr::select(group1,group2,statistic,permutations,dissimilarity,p.val,p.adjust,p.signif) %>%
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
#Jaccard ANOSIM plot
#jaccard_plot <- autoplot(ano, notch =F) + theme_bw() + labs(title = "ANOSIM Jaccard") + stat_pvalue_manual(jaccard_pair, label = "p.signif.box", y.position = max(ano$dis.rank)-(.5*median(res$dis.rank)), step.increase = .03, tip.length = 0.01, hide.ns=T) + xlab(as.character(ioi)) + ylab("Rank Value")
#jaccard_plot
#save to disk 
#ggsave("ANOSIM_jaccard_boxplot.png", plot=jaccard_plot, path = outdir)
#Jaccard ANOSIM plot No Bracket
jaccard_plot_no_brac <- autoplot(ano, notch =F) + theme_bw() +aes(fill=ano$class.vec) + labs(title = "ANOSIM Jaccard", fill=ioi) +  xlab(as.character(ioi)) + ylab("Rank Value")
jaccard_plot_no_brac
#save to disk 
ggsave("ANOSIM_jaccard_boxplot_no_bracket.png", plot=jaccard_plot_no_brac, path = outdir)

```




### 12.6 Multi-response permutation procedure (MRPP)

```{r Methods to do pairwise MRPP comparisons, echo=FALSE, message=FALSE, warning=FALSE}
pairwise_MRPP <- function(dat, column_of_int, dissim){
  all_comparisons <- unique(sample_data(dat)[[column_of_int]])
  combos <- (t(combn(all_comparisons,2)))
  p_res <- c()
  
  for(i in 1:dim(combos)[1]){
    comp <- prune_samples(sample_data(dat)[[ioi]] %in% combos[i,], dat)
    grp <- get_variable(comp, column_of_int)
  
    mrpp_res <- mrpp(distance(comp, dissim), grp)
    mrpp_stats <- c(as.character(combos[i,1]), as.character(combos[i,2]), round(mrpp_res$delta,6), mrpp_res$permutations, mrpp_res$distance, round(mrpp_res$Pvalue,3))
    #tmp <- c(ano, as.character(item))
    p_res <- rbind(p_res, mrpp_stats)
  
  }
  tmp_2 <- data.frame(p_res)
  colnames(tmp_2)<- c("group1", "group2", "delta", "permutations", "dissimilarity", "p.val")
  result <- adjust_pvalue(tmp_2, p.col = "p.val", output.col = "p.adjust",method = "BH")
  result <- add_significance(result, p.col = "p.adjust", output.col = "p.signif",symbols =c("****", "***", "**", "$*$", "ns"))
  res_df <- data.frame(result)
  res_df <- res_df[mixedorder(res_df$group1),]
  return(res_df)
  
}

#out <- pairwise_anosim(cycle_1, ioi , dissim = "bray")
#out

pairwise_MRPP_uni <- function(dat, column_of_int, weight){
  all_comparisons <- unique(sample_data(dat)[[column_of_int]])
  combos <- (t(combn(all_comparisons,2)))
  p_res <- c()
  
  for(i in 1:dim(combos)[1]){
    comp <- prune_samples(sample_data(dat)[[ioi]] %in% combos[i,], dat)
    grp <- get_variable(comp, column_of_int)
  
    tmp <- UniFrac(comp, weighted = weight)
    mrpp_res <- mrpp(tmp, grp)
    mrpp_stats <- c(as.character(combos[i,1]), as.character(combos[i,2]), round(mrpp_res$delta,6), mrpp_res$permutations, round(mrpp_res$Pvalue,3))
    #tmp <- c(ano, as.character(item))
    p_res <- rbind(p_res, mrpp_stats)
  
  }
  tmp_2 <- data.frame(p_res)
  colnames(tmp_2)<- c("group1","group2", "delta", "permutations", "p")
  result <- adjust_pvalue(tmp_2, p.col = "p", output.col = "p.adj",method = "BH")
  result <- add_significance(result, p.col="p.adj",output.col = "p.signif",symbols =c("****", "***", "**", "$*$", "ns"))
  res_df <- data.frame(result)
  res_df <- res_df[mixedorder(res_df$group1),]
  return(res_df)
  
}
```


```{r MRPP analysis, echo=FALSE, message=FALSE, warning=FALSE}
cycle_1 <- qza_to_phyloseq(table_dada2,rooted_tree,taxonomy_file,metadata_file)
sample_data(cycle_1)[[ioi]] <- factor(sample_data(cycle_1)[[ioi]], levels=ioi_ord)

# Weighted Unifrac
w_uni <- UniFrac(cycle_1, weighted = T)
print("MRPP based on Weighted Unifrac distance")
grp <- get_variable(cycle_1, ioi)
mrpp(w_uni, grouping =  grp)
#MRPP pairwise Weighted Unifrac
w_uni_res <- pairwise_MRPP_uni(cycle_1, ioi, T)
kable(w_uni_res) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))


# Unweighted Unifrac
u_uni <- UniFrac(cycle_1, weighted = F)
grp <- get_variable(cycle_1, ioi)
print("MRPP based on Unweighted Unifrac distance")
mrpp(u_uni, grouping = grp)
# MRPP pairwise Unweighted Unifrac
u_uni_res <- pairwise_MRPP_uni(cycle_1, ioi, F)
kable(u_uni_res) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

# Bray Curtis
print("MRPP based on Bray-Curtis distance")
grp <- get_variable(cycle_1, ioi)
mrpp(distance(cycle_1, "bray"), grouping =  grp)
# MRPP pairwise Bray
bray_res <- pairwise_MRPP(cycle_1, ioi, "bray")
kable(bray_res) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

jaccard <- qiime2R::read_qza("./qiime/core-metric-results/jaccard_distance_matrix.qza")
# Jaccard 
print("MRPP based on Jaccard distance")
grp <- get_variable(cycle_1, ioi)
mrpp(distance(cycle_1, "jaccard"), grouping =  grp)
# MRPP pairwise Jaccard
jaccard_res <- pairwise_MRPP(cycle_1, ioi, "jaccard")
kable(jaccard_res) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))


```




### 12.7 AMOVA

```{r Methods to do pairwise AMOVA comparisons, echo=FALSE, message=FALSE, warning=FALSE}
pairwise_AMOVA <- function(dat, column_of_int, dissim){
  all_comparisons <- unique(sample_data(dat)[[column_of_int]])
  combos <- (t(combn(all_comparisons,2)))
  p_res <- c()
  
  for(i in 1:dim(combos)[1]){
    comp <- prune_samples(sample_data(dat)[[ioi]] %in% combos[i,], dat)
    grp <- get_variable(comp, column_of_int)
    dist <- distance(comp, dissim)
    amova_res <- pegas::amova(as.formula(dist ~ grp))
    amova_stats <- c(as.character(combos[i,1]), as.character(combos[i,2]), round(amova_res$tab$MSD[1],6), dissim, round(amova_res$varcomp$P.value[1],3))
  
    p_res <- rbind(p_res, amova_stats)
  
  }
  tmp_2 <- data.frame(p_res)
  colnames(tmp_2)<- c("group1", "group2", "MSD","distance", "p.val")
  result <- adjust_pvalue(tmp_2, p.col = "p.val", output.col = "p.adjust",method = "BH")
  result <- add_significance(result, p.col = "p.adjust", output.col = "p.signif",symbols =c("****", "***", "**", "$*$", "ns"))
  res_df <- data.frame(result)
  res_df <- res_df[mixedorder(res_df$group1),]
  return(res_df)
  
}

#out <- pairwise_anosim(cycle_1, ioi , dissim = "bray")
#out

pairwise_AMOVA_uni <- function(dat, column_of_int, weight){
  all_comparisons <- unique(sample_data(dat)[[column_of_int]])
  combos <- (t(combn(all_comparisons,2)))
  p_res <- c()
  
  for(i in 1:dim(combos)[1]){
    comp <- prune_samples(sample_data(dat)[[ioi]] %in% combos[i,], dat)
    grp <- get_variable(comp, column_of_int)
  
    tmp <- UniFrac(comp, weighted = weight)
    amova_res <- pegas::amova(as.formula(tmp ~ grp))
    amova_stats <- c(as.character(combos[i,1]), as.character(combos[i,2]),weight, round(amova_res$tab$MSD[1],6),  round(amova_res$varcomp$P.value[1],3))
 
    p_res <- rbind(p_res, amova_stats)
  
  }
  tmp_2 <- data.frame(p_res)
  colnames(tmp_2)<- c("group1", "group2", "MSD","weighted", "p.val")
  result <- adjust_pvalue(tmp_2, p.col = "p.val", output.col = "p.adj",method = "BH")
  result <- add_significance(result, p.col="p.adj",output.col = "p.signif",symbols =c("****", "***", "**", "$*$", "ns"))
  res_df <- data.frame(result)
  res_df <- res_df[mixedorder(res_df$group1),]
  return(res_df)
  
}
```


```{r AMOVA testing, echo=FALSE, message=FALSE, warning=FALSE}
cycle_1 <- qza_to_phyloseq(table_dada2,rooted_tree,taxonomy_file,metadata_file)
sample_data(cycle_1)[[ioi]] <- factor(sample_data(cycle_1)[[ioi]], levels=ioi_ord)

#AMOVA Weighted Unifrac
tmp <- UniFrac(cycle_1, weighted = T)
w_uni_grp <- get_variable(cycle_1, ioi)
res <- pegas::amova(tmp ~ w_uni_grp)
res
#AMOVA Weighted unifrac pairwise
w_uni_pair <- pairwise_AMOVA_uni(cycle_1, ioi, T)
kable(w_uni_pair) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

#AMOVA Unweighted Unifrac
tmp <- UniFrac(cycle_1, weighted=F)
u_uni_grp <- get_variable(cycle_1, ioi)
res <- pegas::amova(tmp ~ u_uni_grp)
res
#AMOVA Unweighted Unifrac pairwise
u_uni_pair <- pairwise_AMOVA_uni(cycle_1, ioi, F)
kable(u_uni_pair) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))


#AMOVA bray curtis
dist <- distance(cycle_1, 'bray')
bray_grp <- get_variable(cycle_1, ioi)
res <- pegas::amova(as.formula(dist ~ bray_grp))
res
#AMOVA bray curtis pairwise
bray_amova_pairwise <- pairwise_AMOVA(cycle_1, ioi, 'bray')
kable(bray_amova_pairwise) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

#AMOVA Jaccard 
dist <- distance(cycle_1, 'jaccard')
jaccard_grp <- get_variable(cycle_1, ioi)
res <- pegas::amova(dist ~ jaccard_grp)
res
#AMOVA Jaccard Pairwise 
jaccard_amova_pairwise <- pairwise_AMOVA(cycle_1, ioi, 'jaccard')
kable(jaccard_amova_pairwise) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```




## 13 Lefse

```{r echo=FALSE, message=FALSE, warning=FALSE}
if(getOutputFormat() =='pdf_document'){
  print("see end of report")
}

```

```{r Lefse analysis Results, echo=FALSE, message=FALSE, warning=FALSE, fig.height=25, fig.width=10, out.height="85%", out.width="85%", fig.cap="\\newline{}"}

if(getOutputFormat() == 'html_document') {

  lefse_res_filenames <- list.files("lefse/result",pattern="*_res.png",full.names = T)

  lefse_filenames <- mixedsort(lefse_res_filenames)

  include_graphics(lefse_filenames)
  
} 
```


```{r Lefse analysis Results pdf style, echo=FALSE, message=FALSE, warning=FALSE, out.height="85%",dpi=36, out.width="85%", fig.cap="\\newline{}"}
if(getOutputFormat() =='pdf_document'){
  print("see end of report")
   
  lefse_res_filenames <- list.files("lefse/result",pattern="*_pdf_r.png",full.names=T)

  lefse_filenames <- mixedsort(lefse_res_filenames)
  
  include_graphics(lefse_filenames)
 
}

```


```{r Lefse cladograms, echo=FALSE, message=FALSE, warning=FALSE, fig.height=25, fig.width=10, fig.pos='h', out.width="100%", fig.cap="\\newline{}"}

if(getOutputFormat() == 'html_document') {
 
  
  lefse_clado_filenames <- list.files("lefse/result",pattern="*_cladogram.png",full.names=T)
  
  lefse_filenames <- mixedsort(lefse_clado_filenames)

  include_graphics(lefse_filenames)
  
} 
 
if(getOutputFormat() =='pdf_document'){
  print("see end of report")
  
  lefse_clado_filenames <- list.files("lefse/result",pattern="result/*_pdf_qual_clad.png",full.names=T)
  
  lefse_filenames <- mixedsort(lefse_clado_filenames)

  include_graphics(lefse_filenames)
  
}

```


## 14 References

Report Generation Resources

R Core Team (2020). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria.URL https://www.R-project.org/.

Yihui Xie (2021). knitr: A General-Purpose Package for Dynamic Report Generation in R. R package version 1.31.

\

[1 Stacked Bar Plot References][1 Species Distribution]

---

* Bolyen, E., Rideout, J.R., Dillon, M.R. et al. Reproducible, interactive, scalable and extensible microbiome data science using QIIME 2. Nat Biotechnol 37, 852857 (2019). https://doi.org/10.1038/s41587-019-0209-9

* qiime2R: Importing QIIME2 artifacts and associated data into R sessions. Jordan E Bisanz (2018) https://github.com/jbisanz/qiime2R.

* phyloseq: An R package for reproducible interactive analysis and graphics of microbiome census data. Paul J. McMurdie and Susan Holmes (2013) PLoS ONE 8(4):e61217. 

* Hadley Wickham (2021). forcats: Tools for Working with Categorical Variables (Factors). R package version 0.5.1.
https://CRAN.R-project.org/package=forcats

* H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, 2016.

\

[2 Phylogenetic trees][2 Phylogentic Trees]

---

* export2graphlan.py (ver. 0.2.1 of 27 October 2018). Convert MetaPhlAn, LEfSe, and/or HUMAnN output to GraPhlAn input format. Authors: Francesco Asnicar 
(f.asnicar@unitn.it)

* Asnicar F, Weingart G, Tickle TL, Huttenhower C, Segata N. 2015. Compact graphical representation of phylogenetic data and metadata with GraPhlAn. PeerJ 3:e1029 https://doi.org/10.7717/peerj.1029

\

[3 Heatmap Plot][3 Species Abundance Heatmap]

---

* phyloseq_to_ampvis2.R Andersen K.S. (2019), GitHub gist, https://gist.github.com/KasperSkytte/8d0ca4206a66be7ff6d76fc4ab8e66c6 

* ampvis2: an R package to analyse and visualise 16S rRNA amplicon  data. Andersen K.S., Kirkegaard R.H., Karst S.M., Albertsen M. (2018)  bioRxiv. doi: https://doi.org/10.1101/299537

\

[5 Alpha Diversity Plots][5 Alpha Diversity Plot]

---

* Alboukadel Kassambara (2020). rstatix: Pipe-Friendly Framework for Basic Statistical Tests. R package version 0.6.0. https://CRAN.R-project.org/package=rstatix

* Alboukadel Kassambara (2020). ggpubr: 'ggplot2' Based Publication Ready Plots. R package version 0.4.0.  https://CRAN.R-project.org/package=ggpubr

\

[6 Beta Diversity Ordinations][6 Beta diversity]

---

* Charles A. Martin (2017). ggConvexHull: Add a convex hull geom to ggplot2. R package version 0.1.0. http://github.com/cmartin/ggConvexHull

*  Legendre, P. and Gallagher, E.D. (2001). Ecologically meaningful transformations for ordination of
species data. Oecologia 129, 271-280.

\

[7 Rarefaction Curve][7 Rarefaction Curves]

---

* Hadley Wickham (2007). Reshaping Data with the reshape Package. Journal of Statistical Software, 21(12), 1-20. URL http://www.jstatsoft.org/v21/i12/

\

[9 Unifrac Heatmap][9 Unifrac Heatmap]

---

* Raivo Kolde (2019). pheatmap: Pretty Heatmaps. R package version 1.0.12. https://CRAN.R-project.org/package=pheatmap

\

[11 UPGMA Plots][11 UPGMA Plots]

---

* Andrie de Vries and Brian D. Ripley (2020). ggdendro: Create Dendrograms and Tree Diagrams Using 'ggplot2'. R package version 0.1.22. https://CRAN.R-project.org/package=ggdendro

* Sarkar, Deepayan (2008) Lattice: Multivariate Data Visualization with R. Springer, New York. ISBN 978-0-387-75968-5

\

[12 Beta Diversity Statistical Tests][12 Beta Diversity Statistics]

---

* Jari Oksanen, F. Guillaume Blanchet, Michael Friendly, Roeland Kindt, Pierre Legendre, Dan McGlinn, Peter R. Minchin, R. B. O'Hara, Gavin L. Simpson, Peter Solymos, M. Henry H. Stevens, Eduard Szoecs and Helene Wagner (2020). vegan: Community Ecology Package. R package version 2.5-7. https://CRAN.R-project.org/package=vegan

* Kirill Mller and Hadley Wickham (2021). tibble: Simple Data Frames. R package version 3.0.6. https://CRAN.R-project.org/package=tibble

* Gregory R. Warnes, Ben Bolker and Thomas Lumley (2020). gtools: Various R Programming Tools. R package version 3.8.2. https://CRAN.R-project.org/package=gtools

* James M. Ward (2021). jamba: Jam Base Methods. R package version 0.0.61.900. http://github.com/jmw86069/jamba

* Hao Zhu (2020). kableExtra: Construct Complex Table with 'kable' and Pipe Syntax. R package version 1.3.1. https://CRAN.R-project.org/package=kableExtra

* Paradis E. 2010. pegas: an R package for population genetics with an integrated-modular approach. Bioinformatics 26: 419-420.

\

[13 Lefse][13 Lefse]

---

* Segata N, Izard J, Waldron L, Gevers D, Miropolsky L, Garrett WS, Huttenhower C. Metagenomic biomarker discovery and explanation. Genome Biol. 2011 Jun 24;12(6):R60. doi: 10.1186/gb-2011-12-6-r60. PMID: 21702898; PMCID: PMC3218848.

* Hadley Wickham, Romain Franois, Lionel Henry and Kirill Mller (2021). dplyr: A Grammar of Data Manipulation. R package version 1.0.4. https://CRAN.R-project.org/package=dplyr

